---
title: "Analyse WES CNR"
date: "2023-09-12"
output: 
  html_document:
    toc: true
    toc_float:
      collapsed: true
    toc_depth: 3
    number_sections: true
    theme: lumen
---

# Background

In the previous part of the analysis, we extracted copy number data from exome sequencing. Here we plot the data in the form of heatmap and calculate the correlation. 

# Objectives

1. Calculate risk score


# Objective 1 

**Calculate risk score**

Lets generate a cox-proportional model.

## Overal survival

### Liquid data

We first predict the overall survival / PFS

We start by the trying to replicate the old data from Manish 2022 paper. Lets read the overall surival time and outcome.

```{r}
# file_dir <- file.path(wd$data, "from_papers/cancers_2022_plasma/")
# # Visit 1
# os_v1 <- readxl::read_excel(file.path(file_dir, "Table S3.xlsx"),
#                               sheet = 2) %>%
#   select(PatientID...26:Progression) %>%
#   rename(PatientID = PatientID...26,
#          Death = 'Death (0-no death event)',
#          OS = 'OS (Days)',
#          Res = 'Response at 3-months; 0=no response',
#          PFS = 'PFS (Days)',
#          Prog = Progression) %>%
#   # Convert to month
#   # divide by the average day by month.
#   mutate(OS = round(OS/30, digit=2),
#          PFS = round(PFS/30.417, digit=2))
# 
# # Visit 2
# os_v2 <- readxl::read_excel(file.path(file_dir, "Table S3.xlsx"),
#                               sheet = 3) %>%
#   select(PatientID...26:Progression) %>%
#   rename(PatientID = PatientID...26,
#          Death = 'Death (0-no death event)',
#          OS = 'OS (Days)',
#          Res = 'Response at 3-months; 0=no response',
#          PFS = 'PFS (Days)',
#          Prog = Progression) %>%
#   # Convert to month
#   # divide by the average day by month.
#   mutate(OS = round(OS/30.417, digit=2),
#          PFS = round(PFS/30.417, digit=2))
```

Read the CNR results and join with the metadata above. This is for visit 1


```{r}
file_dir <- file.path(wd$data, "from_papers/cancers_2022_plasma/")
cnr_v1 <- readxl::read_excel(file.path(file_dir, "Table S3.xlsx"),
                             sheet = 2)  %>%
  select(PatientID...1, AR, COL22A1, MYC, NCOR1, NKX3.1, NOTCH1, PIK3CA, PIK3CB, TP53, ZBTB16, TMPRSS2) %>%
  mutate(across(AR:TMPRSS2, ~ case_when(
    . == "NC" ~ 0,
    . == "Amp" ~ 1,
    . == "Del" ~ -1))) %>% as.data.frame() %>%
  rename(PatientID = PatientID...1) %>% 
  mutate(across(AR:TMPRSS2, ~ abs(.)))

df_join_v1 <- cnr_v1 %>% left_join(os_v1) # --- Swap visit 1 or 2

```

Run Cox proportional hazard models for one gene

```{r}
# Fit one gene
tmp_df <- df_join_v1 %>% select(AR:TMPRSS2, OS, Death) # ------- CAHNGE TO OS = Death OR PFS - Prog
cox_model2 <- coxph(Surv(OS, Death) ~ AR, data = tmp_df)
cox_model3 <- coxph(Surv(OS, Death) ~ ., data = tmp_df)

cox_model2
print("---------------")
cox_model3
```

The results are different in each analysis, not sure why. For now we stick to run the gene 1 by 1. Lets run the loop.


```{r}
cox_results <- list()
gene_col <- df_join_v1 %>% select(AR:TMPRSS2) %>% colnames()
for (gene in colnames(df_join_v1)[2:12]) {
  
  tmp_df <- df_join_v1 %>% select(!!as.name(gene), OS, Death) %>%  # ------- CAHNGE TO OS = Death OR PFS - Prog
    rename(gene = !!as.name(gene))
  cox_model <- summary(coxph(Surv(OS, Death) ~ gene, data = tmp_df))
  
  cox_results[[gene]] <- cox_model$coefficients
}

cox_result_v1 <- do.call(rbind, cox_results) %>% as.data.frame()
rownames(cox_result_v1) <- gene_col

df_cox_v1 <- tibble(rownames_to_column(cox_result_v1, var = "Gene")) %>%
  rename(pVal = 'Pr(>|z|)')

```


Now lets calculate the risk score. According to the manuscript

> Sum [Cox regression coefficient × CNA status (1 for gain or loss, and 0 for no change) of each gene selected (ranked by best p values)].


```{r}
# So for a patient we can do like this
patID <- "MJ1"
pat_1 <- df_join_v1 %>% filter(PatientID == patID) %>% select(PatientID, AR:TMPRSS2) %>% 
  pivot_longer(!PatientID, names_to = "Gene", values_to = "CNR") %>%
  left_join(df_cox_v1) %>%
  # Calculate risk score
  mutate(RS = coef * CNR)
pat_sum <- sum(pat_1$RS)
```

Repeat for all patient


```{r,warning=FALSE, message=FALSE}
pat_all <- df_join_v1$PatientID
pat_sum_list <- list()

for (i in seq_along(pat_all)){
  patID <- pat_all[i]
  pat_1 <- df_join_v1 %>% filter(PatientID == patID) %>% select(PatientID, AR:TMPRSS2) %>% 
    pivot_longer(!PatientID, names_to = "Gene", values_to = "CNR") %>%
    left_join(df_cox_v1) %>%
    # Calculate risk score
    mutate(RS = coef * CNR)
  # get the sum
  pat_sum_list[[i]] <- sum(pat_1$RS)
}

df_pat_sum <- data.frame(
  PatientID = pat_all,
  Sum_RS = unlist(pat_sum_list)
)
```


Divide the risk score based on median value

```{r}
df_pat_sum <- df_pat_sum %>% 
  mutate(RS_binr = case_when(Sum_RS > median(Sum_RS) ~ "high",
                             Sum_RS <= median(Sum_RS) ~ "low"))
```


Calculate and plot OS

```{r}
df_risk_v1 <- df_join_v1 %>% left_join(df_pat_sum)

# Fit survival data using the Kaplan-Meier method
surv_object <- Surv(time = df_risk_v1$OS, event = df_risk_v1$Death)  # ------- CAHNGE TO OS = Death OR PFS - Prog
# Fit Kaplan-Meier curve
fit1 <- survfit(surv_object ~ RS_binr, data = df_risk_v1)

p_l_OS <- 
  ggsurvplot(fit1, data = df_risk_v1, pval = TRUE, risk.table=TRUE,
             palette = c("#e41a1c", "#7F7F7F"))

pdf(file.path(wd$output, "OS_PROMOTE_liquid.pdf"), 
    width = 12, height = 8)
p_l_OS
dev.off()
```


Save variable

```{r}
df_risk_v1_OS <- df_risk_v1
```


We subset to samples with solid information

```{r,eval=FALSE}
df_risk_v1 <- df_risk_v1_OS %>% 
  filter(PatientID %in% df_risk_v1_OS_solid$PatientID)

# Fit survival data using the Kaplan-Meier method
surv_object <- Surv(time = df_risk_v1$OS, event = df_risk_v1$Death)  # ------- CAHNGE TO OS = Death OR PFS - Prog
# Fit Kaplan-Meier curve
fit1 <- survfit(surv_object ~ RS_binr, data = df_risk_v1)

p_l_OS <- 
  ggsurvplot(fit1, data = df_risk_v1, pval = TRUE, risk.table=TRUE,
             palette = c("#e41a1c", "#7F7F7F"))

pdf(file.path(wd$output, "OS_PROMOTE_liquid_sameSolidPatients.pdf"), 
    width = 12, height = 8)
p_l_OS
dev.off()
```


Now repeat for PFS

```{r, warning=FALSE, message=FALSE}
cox_results <- list()
gene_col <- df_join_v1 %>% select(AR:TMPRSS2) %>% colnames()
for (gene in colnames(df_join_v1)[2:12]) {
  
  tmp_df <- df_join_v1 %>% select(!!as.name(gene), PFS, Prog) %>%  # ------- CAHNGE TO OS = Death OR PFS - Prog
    rename(gene = !!as.name(gene))
  cox_model <- summary(coxph(Surv(PFS, Prog) ~ gene, data = tmp_df))
  
  cox_results[[gene]] <- cox_model$coefficients
}

cox_result_v1 <- do.call(rbind, cox_results) %>% as.data.frame()
rownames(cox_result_v1) <- gene_col

df_cox_v1 <- tibble(rownames_to_column(cox_result_v1, var = "Gene")) %>%
  rename(pVal = 'Pr(>|z|)')

# Calculate risk score

pat_all <- df_join_v1$PatientID
pat_sum_list <- list()

for (i in seq_along(pat_all)){
  patID <- pat_all[i]
  pat_1 <- df_join_v1 %>% filter(PatientID == patID) %>% select(PatientID, AR:TMPRSS2) %>% 
    pivot_longer(!PatientID, names_to = "Gene", values_to = "CNR") %>%
    left_join(df_cox_v1) %>%
    # Calculate risk score
    mutate(RS = coef * CNR)
  # get the sum
  pat_sum_list[[i]] <- sum(pat_1$RS)
}

df_pat_sum <- data.frame(
  PatientID = pat_all,
  Sum_RS = unlist(pat_sum_list)
)

df_pat_sum <- df_pat_sum %>% 
  mutate(RS_binr = case_when(Sum_RS > median(Sum_RS) ~ "high",
                             Sum_RS <= median(Sum_RS) ~ "low"))

# Fit survival data
df_risk_v1 <- df_join_v1 %>% left_join(df_pat_sum)

# Fit survival data using the Kaplan-Meier method
surv_object <- Surv(time = df_risk_v1$PFS, event = df_risk_v1$Prog)  # ------- CAHNGE TO OS = Death OR PFS - Prog
# Fit Kaplan-Meier curve
fit1 <- survfit(surv_object ~ RS_binr, data = df_risk_v1)

ggsurvplot(fit1, data = df_risk_v1, pval = TRUE, risk.table=TRUE)
```

Save variable

```{r}
df_risk_v1_PFS <- df_risk_v1
```

We subset to samples with solid information

```{r,eval=FALSE}
df_risk_v1 <- df_risk_v1_PFS %>% 
  filter(PatientID %in% df_risk_v1_OS_solid$PatientID)

# Fit survival data using the Kaplan-Meier method
surv_object <- Surv(time = df_risk_v1$PFS, event = df_risk_v1$Prog)  # ------- CAHNGE TO OS = Death OR PFS - Prog
# Fit Kaplan-Meier curve
fit1 <- survfit(surv_object ~ RS_binr, data = df_risk_v1)

p_l_PFS <- 
  ggsurvplot(fit1, data = df_risk_v1, pval = TRUE, risk.table=TRUE,
             palette = c("#e41a1c", "#7F7F7F"))

pdf(file.path(wd$output, "PFS_PROMOTE_liquid_sameSolidPatients.pdf"), 
    width = 12, height = 8)
p_l_PFS
dev.off()
```

#### Visit 2

Repeat for visit 2 OS

```{r, warning=FALSE, message=FALSE}

cnr_v2 <- readxl::read_excel(file.path(file_dir, "Table S3.xlsx"),
                             sheet = 3)  %>%
  select(PatientID...1, AR, COL22A1, MYC, NCOR1, NKX3.1, NOTCH1, PIK3CA, PIK3CB, TP53, ZBTB16, TMPRSS2) %>%
  mutate(across(AR:TMPRSS2, ~ case_when(
    . == "NC" ~ 0,
    . == "Amp" ~ 1,
    . == "Del" ~ -1))) %>% as.data.frame() %>%
  rename(PatientID = PatientID...1) %>% 
  mutate(across(AR:TMPRSS2, ~ abs(.)))

df_join_v1 <- cnr_v2 %>% left_join(os_v2) # --- Swap visit 1 or 2

cox_results <- list()
gene_col <- df_join_v1 %>% select(AR:TMPRSS2) %>% colnames()
for (gene in colnames(df_join_v1)[2:12]) {
  
  tmp_df <- df_join_v1 %>% select(!!as.name(gene), OS, Death) %>%  # ------- CAHNGE TO OS = Death OR PFS - Prog
    rename(gene = !!as.name(gene))
  cox_model <- summary(coxph(Surv(OS, Death) ~ gene, data = tmp_df))
  
  cox_results[[gene]] <- cox_model$coefficients
}

cox_result_v1 <- do.call(rbind, cox_results) %>% as.data.frame()
rownames(cox_result_v1) <- gene_col

df_cox_v1 <- tibble(rownames_to_column(cox_result_v1, var = "Gene")) %>%
  rename(pVal = 'Pr(>|z|)')

# Calculate risk score

pat_all <- df_join_v1$PatientID
pat_sum_list <- list()

for (i in seq_along(pat_all)){
  patID <- pat_all[i]
  pat_1 <- df_join_v1 %>% filter(PatientID == patID) %>% select(PatientID, AR:TMPRSS2) %>% 
    pivot_longer(!PatientID, names_to = "Gene", values_to = "CNR") %>%
    left_join(df_cox_v1) %>%
    # Calculate risk score
    mutate(RS = coef * CNR)
  # get the sum
  pat_sum_list[[i]] <- sum(pat_1$RS)
}

df_pat_sum <- data.frame(
  PatientID = pat_all,
  Sum_RS = unlist(pat_sum_list)
)

df_pat_sum <- df_pat_sum %>% 
  mutate(RS_binr = case_when(Sum_RS > median(Sum_RS) ~ "high",
                             Sum_RS <= median(Sum_RS) ~ "low"))

# Fit survival data
df_risk_v1 <- df_join_v1 %>% left_join(df_pat_sum)

# Fit survival data using the Kaplan-Meier method
surv_object <- Surv(time = df_risk_v1$OS, event = df_risk_v1$Death)  # ------- CAHNGE TO OS = Death OR PFS - Prog
# Fit Kaplan-Meier curve
fit1 <- survfit(surv_object ~ RS_binr, data = df_risk_v1)

ggsurvplot(fit1, data = df_risk_v1, pval = TRUE, risk.table=TRUE)
```


Save variable

```{r}
df_risk_v2_OS <- df_risk_v1
```


### Solid data

#### Visit 1

We established that we can replicate the previous analysis, now lets run the same analysis on solid (exome) data

Step 1 - Combine the CNR solid call with survival data

```{r}
# WITHOUT FILTERING
cnr_solid_v1 <- abs(mat_solid_v1) %>% as.data.frame() %>% t() # -- without filtering


# # FILTER TO SPECIFIC SITE
# dd <- meta_v1 %>% left_join(meta_l)
# # Filter to only bone site
# bone_samp <- dd %>% filter(Sample_type == "Bone") %>% filter(Type == "Promote1") %>% pull(PatientID)
# bone_samp <- bone_samp[bone_samp %in% colnames(mat_v1_or2)]
# mat_v1_sub <- mat_v1_or2[,bone_samp]

# Visit 2
#bone_samp <- dd %>% filter(Sample_type == "non-bone") %>% filter(Type == "Promote2") %>% pull(PatientID)
#bone_samp <- bone_samp[bone_samp %in% colnames(mat_v2_or2)]
#mat_v1_sub <- mat_v2_or2[,bone_samp] # Visit 2


#cnr_solid_v1 <- abs(mat_v1_sub) %>% as.data.frame() %>% t() 

#cnr_solid_v1 <- abs(pap_v1) %>% as.data.frame() %>% t()
colnames(cnr_solid_v1) <- gsub(" .*", "", colnames(cnr_solid_v1)) 
cnr_solid_v1 <- cnr_solid_v1 %>% as.data.frame() %>%
  tibble::rownames_to_column(var = "PatientID") %>% 
  left_join(os_v1) %>%        # EDIT FOR VISIT 1 or VISIT 2
  rename(NKX3.1 = 'NKX3-1')

# add additionall annotation
df_purity_sub <- df_purity_or_v1 %>% select(PatientID, Purity_tresh)
cnr_solid_v1 <- cnr_solid_v1 %>% left_join(df_purity_sub) #%>%
#filter(!is.na(Purity_tresh))
```


Step 2 - Calculate cox-regression based on PFS


```{r}
cox_results <- list()
gene_col <- cnr_solid_v1 %>% select(AR:NKX3.1 ) %>% colnames()
for (gene in colnames(cnr_solid_v1)[2:12]) {
  
  tmp_df <- cnr_solid_v1 %>% select(!!as.name(gene), PFS, Prog) %>% # ------- CAHNGE TO OS = Death OR PFS - Prog
    rename(gene = !!as.name(gene))
  cox_model <- summary(coxph(Surv(PFS, Prog) ~ gene, data = tmp_df))
  
  cox_results[[gene]] <- cox_model$coefficients
}

cox_result_v1 <- do.call(rbind, cox_results) %>% as.data.frame()
rownames(cox_result_v1) <- gene_col

df_cox_v1 <- tibble(rownames_to_column(cox_result_v1, var = "Gene")) %>%
  rename(pVal = 'Pr(>|z|)') %>%
  filter(!is.na(coef))
```



Step 3 - Calculate risk score


```{r,warning=FALSE, message=FALSE}
# Only for bone / non-bone
#pat_all <- bone_samp

# All patient
pat_all <- cnr_solid_v1$PatientID
pat_sum_list <- list()

for (i in seq_along(pat_all)){
  patID <- pat_all[i]
  pat_1 <- cnr_solid_v1 %>% filter(PatientID == patID) %>% select(PatientID, df_cox_v1$Gene) %>% 
    pivot_longer(!PatientID, names_to = "Gene", values_to = "CNR") %>%
    left_join(df_cox_v1) %>%
    # Calculate risk score
    mutate(RS = coef * CNR)
  # get the sum
  pat_sum_list[[i]] <- sum(pat_1$RS)
}

df_pat_sum <- data.frame(
  PatientID = pat_all,
  Sum_RS = unlist(pat_sum_list)
)


df_pat_sum <- df_pat_sum %>% 
  mutate(RS_binr = case_when(Sum_RS > median(Sum_RS) ~ "high",
                             Sum_RS <= median(Sum_RS) ~ "low"))

# Calculate and plot PFS
df_risk_solid_v1 <- cnr_solid_v1 %>% left_join(df_pat_sum)

# Fit survival data using the Kaplan-Meier method
surv_object <- Surv(time = df_risk_solid_v1$PFS, event = df_risk_solid_v1$Prog) # ------- CAHNGE TO OS = Death OR PFS - Prog
# Fit Kaplan-Meier curve
fit1 <- survfit(surv_object ~ RS_binr, data = df_risk_solid_v1)
p_PFS <- ggsurvplot(fit1, data = df_risk_solid_v1, pval = TRUE, risk.table=TRUE)
p_PFS

# # add additionall annotation
# kk <- df_purity_or %>% select(PatientID, Body_Site)
# df_risk_solid_v1 <- df_risk_solid_v1 %>% left_join(df_purity_sub) %>%
#   left_join(kk) %>%
#   mutate(Cat = paste(RS_binr, Purity_tresh, sep="_"),
#          Cat2 = case_when(Cat == "high_t1_high" ~ "High.risk_High.Purity",
#                           TRUE ~ "Low.risk"),
#          Cat3 = paste(Body_Site, Purity_tresh, sep="_")) 

# # If we want to stratify, re-calculate
# df_risk_solid_v1_sub <- df_risk_solid_v1 %>% filter(!is.na(Purity_tresh)) %>%
#   # Remove the low number of samples from the low purity
#   #filter(!Cat %in% c("high_t2_low", "low_t2_low"))
#   #filter(Cat %in% c("high_t1_high", "low_t2_low"))
# surv_object2 <- Surv(time = df_risk_solid_v1_sub$OS, event = df_risk_solid_v1_sub$Death) # ------- CAHNGE TO OS = Death OR PFS - Prog
# # Fit Kaplan-Meier curve
# fit2 <- survfit(surv_object2 ~ Cat, data = df_risk_solid_v1_sub)
# p_sur <- ggsurvplot(fit2, data = df_risk_solid_v1_sub, pval = TRUE, risk.table=TRUE,
#                     palette = c("red", "pink", "darkblue", "lightblue"))
# p_sur
```

Save variable

```{r}
df_risk_v1_PFS_solid <- df_risk_solid_v1
```

Make nicer figure and save

```{r}
# Fit survival data using the Kaplan-Meier method
surv_object <- Surv(time = df_risk_v1_PFS_solid$PFS, event = df_risk_v1_PFS_solid$Prog)  # ------- CAHNGE TO OS = Death OR PFS - Prog
# Fit Kaplan-Meier curve
fit1 <- survfit(surv_object ~ RS_binr, data = df_risk_v1_PFS_solid)

pdf(file.path(wd$output, "PFS_visit_1.pdf"), width = 12, height = 8)
ggsurvplot(fit1, data = df_risk_v1_PFS_solid, pval = TRUE, risk.table=TRUE,
           conf.int = TRUE,
           palette = c("red", "grey50"))
dev.off()
```

Repeat for OS

```{r, warning=FALSE, message=FALSE}
# Calculate cox

cox_results <- list()
gene_col <- cnr_solid_v1 %>% select(AR:NKX3.1 ) %>% colnames()
for (gene in colnames(cnr_solid_v1)[2:12]) {
  
  tmp_df <- cnr_solid_v1 %>% select(!!as.name(gene), OS, Death) %>% # ------- CAHNGE TO OS = Death OR PFS - Prog
    rename(gene = !!as.name(gene))
  cox_model <- summary(coxph(Surv(OS, Death) ~ gene, data = tmp_df))
  
  cox_results[[gene]] <- cox_model$coefficients
}

cox_result_v1 <- do.call(rbind, cox_results) %>% as.data.frame()
rownames(cox_result_v1) <- gene_col

df_cox_v1 <- tibble(rownames_to_column(cox_result_v1, var = "Gene")) %>%
  rename(pVal = 'Pr(>|z|)') %>%
  filter(!is.na(coef))

# Save the weignt
df_cox_visit_1 <- df_cox_v1

# Cacluate Risk score
# All patient
pat_all <- cnr_solid_v1$PatientID
pat_sum_list <- list()

for (i in seq_along(pat_all)){
  patID <- pat_all[i]
  pat_1 <- cnr_solid_v1 %>% filter(PatientID == patID) %>% select(PatientID, df_cox_v1$Gene) %>% 
    pivot_longer(!PatientID, names_to = "Gene", values_to = "CNR") %>%
    left_join(df_cox_v1) %>%
    # Calculate risk score
    mutate(RS = coef * CNR)
  # get the sum
  pat_sum_list[[i]] <- sum(pat_1$RS)
}

df_pat_sum <- data.frame(
  PatientID = pat_all,
  Sum_RS = unlist(pat_sum_list)
)


df_pat_sum <- df_pat_sum %>% 
  mutate(RS_binr = case_when(Sum_RS > median(Sum_RS) ~ "high",
                             Sum_RS <= median(Sum_RS) ~ "low"))

# Calculate and plot OS
df_risk_solid_v1 <- cnr_solid_v1 %>% left_join(df_pat_sum)

# Fit survival data using the Kaplan-Meier method
surv_object <- Surv(time = df_risk_solid_v1$OS, event = df_risk_solid_v1$Death) # ------- CAHNGE TO OS = Death OR PFS - Prog
# Fit Kaplan-Meier curve
fit1 <- survfit(surv_object ~ RS_binr, data = df_risk_solid_v1)
p_OS <- ggsurvplot(fit1, data = df_risk_solid_v1, pval = TRUE, risk.table=TRUE)
p_OS
```

Save variable


```{r}
df_risk_v1_OS_solid <- df_risk_solid_v1
```

Plot nicer

```{r}
# Fit survival data using the Kaplan-Meier method
surv_object <- Surv(time = df_risk_v1_OS_solid$OS, event = df_risk_v1_OS_solid$Death)  # ------- CAHNGE TO OS = Death OR PFS - Prog
# Fit Kaplan-Meier curve
fit1 <- survfit(surv_object ~ RS_binr, data = df_risk_v1_OS_solid)

pdf(file.path(wd$output, "OS_visit_1.pdf"), width = 12, height = 8)
ggsurvplot(fit1, data = df_risk_v1_OS_solid, pval = TRUE, risk.table=TRUE,
           conf.int = TRUE,
           palette = c("red", "grey50"))
dev.off()

```

Get the min and max OS value for the high risk score

```{r}
summary(filter(df_risk_v1_OS_solid, RS_binr == "high") %>% pull(OS))
summary(filter(df_risk_v1_OS_solid, RS_binr == "low") %>% pull(OS))
```
## Comparing liquid and solid RS


```{r}
solid_v1_rs <- df_risk_v1_OS_solid %>% 
  select(PatientID, Sum_RS, RS_binr, OS) %>% 
  rename(Sol_OS_RS = Sum_RS,
         Sol_OS_bin = RS_binr,
         Sol_OS = OS) %>% 
  mutate(Sol_OS_bin2 = paste0("s_", Sol_OS_bin))


liquid_v1_rs <- df_risk_v1_OS %>% 
  select(PatientID, Sum_RS, RS_binr, OS) %>% 
  rename(Liq_OS_RS = Sum_RS,
         Liq_OS_bin = RS_binr,
         Liq_OS = OS)

v1_rs <- left_join(solid_v1_rs, liquid_v1_rs) %>% 
  mutate(Cat_RS = case_when(
    Sol_OS_bin == "high" & Liq_OS_bin == "high" ~ "both_high",
    Sol_OS_bin == "low" & Liq_OS_bin == "low" ~ "both_low",
    Sol_OS_bin == "high" & Liq_OS_bin == "low" ~ "only_solid_h",
    Sol_OS_bin == "low" & Liq_OS_bin == "high" ~ "only_liquid_h"))


# purity_info <- df_tumor_p3 %>% select(PatientID, ct_cut)
# 
# v1_rs <- v1_rs %>% left_join(purity_info)
# 

vector1 <- v1_rs$Liq_OS_bin
vector2 <- v1_rs$Sol_OS_bin
vector2.2 <- v1_rs$Sol_OS_bin2

table(vector1, vector2.2)

kappa_result2 <- cohen.kappa(x=cbind(vector1,vector2))

ggplot(v1_rs, aes(x=Cat_RS, y=Sol_OS)) +
  geom_boxplot(outlier.alpha = 0,  width = 0.4, fatten=3, alpha=0.6, colour="grey50",
                position = position_dodge(width = 0.9)) +
  geom_quasirandom(dodge.width=0.9, alpha=0.5, groupOnX = TRUE) +
  NULL

```


Limit the concordance to the samples with high tumor purity and high ctDNA fraction

```{r}
g5 <- df_purity_or_v1 %>% filter(Sample_type == "Bone" & Purity_tresh == "t1_high" & ct_tresh == "ct1_high")
```


```{r}
v1_rs_sub <- v1_rs %>% 
  filter(PatientID %in% g5$PatientID)
  
  
vector1 <- v1_rs_sub$Liq_OS_bin
vector2 <- v1_rs_sub$Sol_OS_bin
vector2.2 <- v1_rs_sub$Sol_OS_bin2

table(vector1, vector2.2)

kappa_result2 <- cohen.kappa(x=cbind(vector1,vector2))

ggplot(v1_rs_sub, aes(x=Cat_RS, y=Sol_OS)) +
  geom_boxplot(outlier.alpha = 0,  width = 0.4, fatten=3, alpha=0.6, colour="grey50",
                position = position_dodge(width = 0.9)) +
  geom_quasirandom(dodge.width=0.9, alpha=0.5, groupOnX = TRUE) +
  NULL

```

Limit the concordance to high ctDNA fraction samples

```{r}
g6 <- df_purity_or_v1 %>% filter(ct_tresh == "ct1_low")
```


```{r}
v1_rs_sub <- v1_rs %>% 
  filter(PatientID %in% g6$PatientID)
  
  
vector1 <- v1_rs_sub$Liq_OS_bin
vector2 <- v1_rs_sub$Sol_OS_bin
vector2.2 <- v1_rs_sub$Sol_OS_bin2

table(vector1, vector2.2)

kappa_result2 <- cohen.kappa(x=cbind(vector1,vector2))

ggplot(v1_rs_sub, aes(x=Cat_RS, y=Sol_OS)) +
  geom_boxplot(outlier.alpha = 0,  width = 0.4, fatten=3, alpha=0.6, colour="grey50",
                position = position_dodge(width = 0.9)) +
  geom_quasirandom(dodge.width=0.9, alpha=0.5, groupOnX = TRUE) +
  NULL

```



We look at the range of the risk score in solid and liquid


```{r}
# Extract data 
# 1) Risk score values
RS_liq_PFS <- df_risk_v1_PFS$Sum_RS
RS_liq_OS <- df_risk_v1_OS$Sum_RS

RS_sol_PFS <- df_risk_v1_PFS_solid$Sum_RS
RS_sol_OS <- df_risk_v1_OS_solid$Sum_RS

# 2) Risk score binary
RS_liq_PFS_bin <- df_risk_v1_PFS$RS_binr
RS_liq_OS_bin <- df_risk_v1_OS$RS_binr

RS_sol_PFS_bin <- df_risk_v1_PFS_solid$RS_binr
RS_sol_OS_bin <- df_risk_v1_OS_solid$RS_binr

# 3) Patient ID
id_1 <- df_risk_v1_PFS$PatientID
id_2 <- df_risk_v1_OS$PatientID

id_3 <- df_risk_v1_PFS_solid$PatientID
id_4 <- df_risk_v1_OS_solid$PatientID


df_risk_values <-
  data.frame(
    PatientID = c(id_1, id_2, id_3, id_4),
    RiskValue = c(RS_liq_PFS, RS_liq_OS,
                  RS_sol_PFS, RS_sol_OS),
    Source = c(rep("liquid", length(c(RS_liq_PFS, RS_liq_OS))),
               rep("solid", length(c(RS_sol_PFS, RS_sol_OS)))),
    Variable = c(rep("PFS", length(RS_liq_PFS)),
                 rep("OS", length(RS_liq_OS)),
                 rep("PFS", length(RS_sol_PFS)),
                 rep("OS", length(RS_sol_OS))),
    Cat = c(RS_liq_PFS_bin, RS_liq_OS_bin,
                  RS_sol_PFS_bin, RS_sol_OS_bin)
  )

# Plot

ggplot(df_risk_values, aes(x=Source, y=RiskValue, colour=Cat)) +
  #geom_violin(alpha = 0.5, scale = "width", position = position_dodge(width = 0.9), colour="grey50") +
  geom_boxplot(outlier.alpha = 0,  width = 0.1, fatten=3, alpha=0.6, colour="grey50",
                position = position_dodge(width = 0.9)) +
  geom_quasirandom(dodge.width=0.9, alpha=0.5, groupOnX = TRUE) +
  facet_wrap(~Variable, ncol=1) +
  theme_bw() +
  theme(legend.position = "none",
          panel.grid.major = element_line(linetype = "blank"), 
          panel.grid.minor = element_line(linetype = "blank")) +
  scale_colour_manual(values=c("red", "lightblue"))
```


Lets show it in a heatmap style

```{r}
df_1_tmp <- df_risk_v1_PFS %>% select(PatientID, RS_binr) %>% rename(RS_PFS_l = RS_binr)
df_2_tmp <- df_risk_v1_OS %>% select(PatientID, RS_binr) %>% rename(RS_OS_l = RS_binr)
df_3_tmp <- df_risk_v1_PFS_solid %>% select(PatientID, RS_binr) %>% rename(RS_PFS_s = RS_binr)
df_4_tmp <- df_risk_v1_OS_solid %>% select(PatientID, RS_binr) %>% rename(RS_OS_s = RS_binr)

df_tmp <- df_1_tmp %>% left_join(df_2_tmp) %>% left_join(df_3_tmp) %>% left_join(df_4_tmp) %>%
  mutate(across(RS_PFS_l:RS_OS_s, ~ case_when(
    . == "low" ~ 0,
    . == "high" ~ 1))) %>% as.data.frame() %>%
  arrange(desc(RS_PFS_s), desc(RS_OS_s), desc(RS_PFS_l), desc(RS_OS_l))


mm <- melt(df_tmp) %>%
  mutate(value = as.character(value),
         PatientID = factor(PatientID, levels=df_tmp$PatientID))



ggplot(mm, aes(x=PatientID, y=variable, colour=value)) +
  geom_point(size=4, shape=15) +
  theme_bw() +
  theme(legend.position = "bottom",
        panel.grid.major = element_line(linetype = "blank"), 
        panel.grid.minor = element_line(linetype = "blank"),
        axis.text.x = element_blank()) +
  scale_colour_discrete(type=c("lightblue", "red")) +
  NULL

```
Only the OS

```{r}
int_sample <- intersect(df_2_tmp$PatientID, df_4_tmp$PatientID)
mm_OS <- mm %>%  filter(variable %in% c("RS_OS_s", "RS_OS_l")) %>% 
  filter(PatientID %in% int_sample) 

# Add purity info
mm_OS <- mm_OS %>%  left_join(df_purity_or_v1)

# Manually order based on input file
new_or <- read_xlsx(file.path(wd$data, "heatmap_risk_or.xlsx"), sheet = 2) %>% pull(New_or)
  
mm_OS <- mm_OS %>% 
  mutate(PatientID = factor(PatientID, levels=new_or))

ggplot(mm_OS, aes(x=PatientID, y=variable, colour=value)) +
  geom_point(size=4, shape=15) +
  theme_bw() +
  theme(legend.position = "bottom",
        panel.grid.major = element_line(linetype = "blank"), 
        panel.grid.minor = element_line(linetype = "blank"),
        axis.text.x = element_blank()) +
  scale_colour_discrete(type=c("lightblue", "red")) +
  NULL


ggsave(file=file.path(wd$output, "RS_OS_solid_liquid.pdf"), device = "pdf", width = 12, height = 4)
```



Calculate the concordence using kappa

```{r}
v1 <- mm_OS %>% filter(variable == "RS_OS_s")
v2 <- mm_OS %>% filter(variable == "RS_OS_l")
identical(v1$PatientID, v2$PatientID)


kappa_result2 <- cohen.kappa(x=cbind(v1$value,v2$value))
kappa_result2
```

Make it as percentage

```{r}
mm_OS_wide <- mm_OS %>% pivot_wider(names_from = variable, values_from = value)

# Calculate percentage
s1 <- mm_OS_wide %>% filter(RS_OS_s == 1) 
s1_l0 = s1 %>% filter(RS_OS_l == 0)
s1_l1 = s1 %>% filter(RS_OS_l == 1)

# Perc in solid
p1 <- nrow(s1_l1)  / nrow(s1) # Both 1
p2 <- nrow(s1_l0)  / nrow(s1) # Solid 1, liquid 0

s0 <- mm_OS_wide %>% filter(RS_OS_s == 0) 
s0_l0 = s0 %>% filter(RS_OS_l == 0)
s0_l1 = s0 %>% filter(RS_OS_l == 1)

# Perc in solid
p3 <- nrow(s0_l0)  / nrow(s0) * 100 # Both 0
p4 <- nrow(s0_l1)  / nrow(s0) * 100 # Solid 0, liquid 1
```



#### Subset - Purity

Lets plot the OS / PFS  based on purity

```{r}
# OS Based on purity
tmp_surv <- df_risk_v1_OS_solid 
surv_object <- Surv(time = tmp_surv$OS, event = tmp_surv$Death) 
fit1 <- survfit(surv_object ~ Purity_tresh, data = tmp_surv)

pdf(file.path(wd$output, "OS_visit_1_byPurity.pdf"), width = 12, height = 8)
ggsurvplot(fit1, data = tmp_surv, pval = TRUE, risk.table=TRUE,
            palette = c("#e41a1c", "#377eb8"))
dev.off()


# PFS Based on purity
tmp_surv <- df_risk_v1_PFS_solid 
surv_object <- Surv(time = df_risk_v1_OS_solid$PFS, event = df_risk_v1_OS_solid$Prog) 
fit1 <- survfit(surv_object ~ Purity_tresh, data = tmp_surv)

pdf(file.path(wd$output, "PFS_visit_1_byPurity.pdf"), width = 12, height = 8)
ggsurvplot(fit1, data = tmp_surv, pval = TRUE, risk.table=TRUE,
            palette = c("#e41a1c", "#377eb8"))
dev.off()
```


#### Subset - Site

```{r}
# OS Based on site
kk <- df_purity_or_v1 %>% select(PatientID, Body_Site)
tmp_surv <- df_risk_v1_OS_solid %>% left_join(kk) 
#tmp_surv <- tmp_surv %>% filter(Purity_tresh == "t2_low")
surv_object <- Surv(time = tmp_surv$OS, event = tmp_surv$Death) 
fit1 <- survfit(surv_object ~ Body_Site, data = tmp_surv)

pdf(file.path(wd$output, "OS_visit_1_bySite.pdf"), width = 12, height = 8)
ggsurvplot(fit1, data = tmp_surv, pval = TRUE, risk.table=TRUE,
            palette = c("#8CD17D", "#FFBE7D"))
dev.off()


# PFS Based on site
tmp_surv <- df_risk_v1_PFS_solid %>% left_join(kk) 
surv_object <- Surv(time = df_risk_v1_OS_solid$PFS, event = df_risk_v1_OS_solid$Prog) 
fit1 <- survfit(surv_object ~ Body_Site, data = tmp_surv)

pdf(file.path(wd$output, "PFS_visit_1_bySite.pdf"), width = 12, height = 8)
ggsurvplot(fit1, data = tmp_surv, pval = TRUE, risk.table=TRUE,
            palette = c("#8CD17D", "#FFBE7D"))
dev.off()
```




#### Visit 2

Run risk score calculation for visit 2

```{r,warning=FALSE, message=FALSE}
# Get copy number call
cnr_solid_v1 <- abs(mat_solid_v2) %>% as.data.frame() %>% t() # -- without filtering

colnames(cnr_solid_v1) <- gsub(" .*", "", colnames(cnr_solid_v1)) 
cnr_solid_v1 <- cnr_solid_v1 %>% as.data.frame() %>%
  tibble::rownames_to_column(var = "PatientID") %>% 
  left_join(os_v2) %>%        # EDIT FOR VISIT 1 or VISIT 2
  rename(NXK3.1 = 'NKX3-1')

# add additionall annotation
df_purity_sub <- df_purity_or_v2 %>% select(PatientID, Purity_tresh)
cnr_solid_v1 <- cnr_solid_v1 %>% left_join(df_purity_sub) #%>%

# Get cox
cox_results <- list()
gene_col <- cnr_solid_v1 %>% select(AR:NXK3.1 ) %>% colnames()
for (gene in colnames(cnr_solid_v1)[2:12]) {
  
  tmp_df <- cnr_solid_v1 %>% select(!!as.name(gene), OS, Death) %>% # ------- CAHNGE TO OS = Death OR PFS - Prog
    rename(gene = !!as.name(gene))
  cox_model <- summary(coxph(Surv(OS, Death) ~ gene, data = tmp_df))
  
  cox_results[[gene]] <- cox_model$coefficients
}

cox_result_v1 <- do.call(rbind, cox_results) %>% as.data.frame()
rownames(cox_result_v1) <- gene_col

df_cox_v1 <- tibble(rownames_to_column(cox_result_v1, var = "Gene")) %>%
  rename(pVal = 'Pr(>|z|)') %>%
  filter(!is.na(coef))

# Only for bone / non-bone
#pat_all <- bone_samp

# All patient
pat_all <- cnr_solid_v1$PatientID
pat_sum_list <- list()

for (i in seq_along(pat_all)){
  patID <- pat_all[i]
  pat_1 <- cnr_solid_v1 %>% filter(PatientID == patID) %>% select(PatientID, df_cox_v1$Gene) %>% 
    pivot_longer(!PatientID, names_to = "Gene", values_to = "CNR") %>%
    left_join(df_cox_v1) %>%
    # Calculate risk score
    mutate(RS = coef * CNR)
  # get the sum
  pat_sum_list[[i]] <- sum(pat_1$RS)
}

df_pat_sum <- data.frame(
  PatientID = pat_all,
  Sum_RS = unlist(pat_sum_list)
)


df_pat_sum <- df_pat_sum %>% 
  mutate(RS_binr = case_when(Sum_RS > median(Sum_RS) ~ "high",
                             Sum_RS <= median(Sum_RS) ~ "low"))

# Calculate risk score
df_risk_v2_OS_solid <- cnr_solid_v1 %>% left_join(df_pat_sum)
```

Repeat for PFS

```{r, warning=FALSE, message=FALSE}
# Get copy number call
cnr_solid_v1 <- abs(mat_solid_v2) %>% as.data.frame() %>% t() # -- without filtering

colnames(cnr_solid_v1) <- gsub(" .*", "", colnames(cnr_solid_v1)) 
cnr_solid_v1 <- cnr_solid_v1 %>% as.data.frame() %>%
  tibble::rownames_to_column(var = "PatientID") %>% 
  left_join(os_v2) %>%        # EDIT FOR VISIT 1 or VISIT 2
  rename(NXK3.1 = 'NKX3-1')

# add additionall annotation
df_purity_sub <- df_purity_or_v2 %>% select(PatientID, Purity_tresh)
cnr_solid_v1 <- cnr_solid_v1 %>% left_join(df_purity_sub) #%>%

# Get cox
cox_results <- list()
gene_col <- cnr_solid_v1 %>% select(AR:NXK3.1 ) %>% colnames()
for (gene in colnames(cnr_solid_v1)[2:12]) {
  
  tmp_df <- cnr_solid_v1 %>% select(!!as.name(gene), PFS, Prog) %>% # ------- CAHNGE TO OS = Death OR PFS - Prog
    rename(gene = !!as.name(gene))
  cox_model <- summary(coxph(Surv(PFS, Prog) ~ gene, data = tmp_df))
  
  cox_results[[gene]] <- cox_model$coefficients
}

cox_result_v1 <- do.call(rbind, cox_results) %>% as.data.frame()
rownames(cox_result_v1) <- gene_col

df_cox_v1 <- tibble(rownames_to_column(cox_result_v1, var = "Gene")) %>%
  rename(pVal = 'Pr(>|z|)') %>%
  filter(!is.na(coef))

# Only for bone / non-bone
#pat_all <- bone_samp

# All patient
pat_all <- cnr_solid_v1$PatientID
pat_sum_list <- list()

for (i in seq_along(pat_all)){
  patID <- pat_all[i]
  pat_1 <- cnr_solid_v1 %>% filter(PatientID == patID) %>% select(PatientID, df_cox_v1$Gene) %>% 
    pivot_longer(!PatientID, names_to = "Gene", values_to = "CNR") %>%
    left_join(df_cox_v1) %>%
    # Calculate risk score
    mutate(RS = coef * CNR)
  # get the sum
  pat_sum_list[[i]] <- sum(pat_1$RS)
}

df_pat_sum <- data.frame(
  PatientID = pat_all,
  Sum_RS = unlist(pat_sum_list)
)


df_pat_sum <- df_pat_sum %>% 
  mutate(RS_binr = case_when(Sum_RS > median(Sum_RS) ~ "high",
                             Sum_RS <= median(Sum_RS) ~ "low"))

# Calculate risk score
df_risk_v2_PFS_solid <- cnr_solid_v1 %>% left_join(df_pat_sum)
```




# Objective 2

## Compare CNR score


```{r}
df_CNR <- df_simple2
prog_dat <- d_v %>% select(SubjectID, Prog)
  
#df_CNR <- df_simple2
df_CNR2 <- df_CNR %>% left_join(prog_dat) %>%
  mutate(Prog = case_when(Prog == 0 ~ "Non-responder",
                          Prog == 1 ~ "Responder"))
```

Example plot

```{r}
tt <- df_CNR2 %>% filter(Genes == "AR")
v1_id <- df_CNR2 %>% filter(visit == "V1") %>% pull(SubjectID)
v2_id <- df_CNR2 %>% filter(visit == "V2") %>% pull(SubjectID)
int_id <- intersect(v1_id, v2_id)

tt2 <- filter(tt, SubjectID %in% int_id) %>% distinct()
#write.csv(tt2, "~/Desktop/tmp.csv")

ggplot(tt2, aes(x=visit, y=lg2CNR)) +
  geom_point() +
  geom_line(aes(group=SubjectID)) +
  facet_wrap(Genes~Prog, ncol=6) +
  geom_signif(
    comparisons = list(c("V1", "V2")),
    map_signif_level = TRUE, textsize = 6,
    test = "t.test",
    test.args = list(paired = TRUE),
  ) + theme_bw() +
  NULL
  

```


Now repeat for each plot

```{r}
# Repeat for each plot
df_CNR3 <- df_CNR2 %>% filter(SubjectID %in% int_id) %>% distinct()

list_p <- list()

uq <- unique(df_CNR3$Genes)

for (i in seq_along(uq)) {
  
  tt2 <- df_CNR3 %>% filter(Genes == uq[i])
  
  # Get 20% of max to increase scale to add the significant bar
  min_val <- min(tt2$lg2CNR)
  max_val <- max(tt2$lg2CNR)
  max_val <- max_val * 2
  
  pp <- ggplot(tt2, aes(x=visit, y=lg2CNR)) +
    geom_point() +
    geom_line(aes(group=SubjectID)) +
    facet_wrap(Genes~Prog, ncol=6) +
    geom_signif(
      comparisons = list(c("V1", "V2")),
      map_signif_level = TRUE, textsize = 6,
      test = "t.test",
      test.args = list(paired = TRUE),
    ) + theme_bw() +
    scale_y_continuous(limits=c(min_val, max_val ))
    NULL
  
  list_p[[i]] <- pp
}

pdf(file.path(wd$output, "CNR_genes_Res_vs_nonRes.pdf"), height = 10, width = 15)
list_p[[1]] + list_p[[2]] +
  list_p[[3]] + list_p[[4]] +
  list_p[[5]] + list_p[[6]] +
  list_p[[7]] + list_p[[8]] +
  list_p[[9]] + list_p[[10]] + 
  list_p[[11]] + plot_layout(ncol=4)
dev.off()

```

## Compare risk score

We compare the risk score of OS and PFS for visit1 and visit 2



```{r}
# df linking patientID with subject ID
tmp_df_1 <- d_v %>% select(SubjectID, Visit_1_ID) %>% rename(PatientID = Visit_1_ID) %>% distinct()
tmp_df_2 <- d_v %>% select(SubjectID, Visit_2_ID) %>% rename(PatientID = Visit_2_ID) %>% distinct()

# OS
t1 <- df_risk_v1_OS_solid %>% mutate(Visit = "V1") %>% select(PatientID, Sum_RS, Visit, Res, RS_binr) %>% left_join(tmp_df_1)
t2 <- df_risk_v2_OS_solid %>% mutate(Visit = "V2") %>% select(PatientID, Sum_RS, Visit, Res, RS_binr) %>% left_join(tmp_df_2)
df_os_risk <- rbind(t1, t2) 

# PFS
t1 <- df_risk_v1_PFS_solid %>% mutate(Visit = "V1") %>% select(PatientID, Sum_RS, Visit, Res, RS_binr) %>% left_join(tmp_df_1)
t2 <- df_risk_v2_PFS_solid %>% mutate(Visit = "V2") %>% select(PatientID, Sum_RS, Visit, Res, RS_binr) %>% left_join(tmp_df_2)
df_pfs_risk <- rbind(t1, t2) 

```

Plot

```{r}
int_id <- intersect(t1$SubjectID, t2$SubjectID)

df_os_risk <- filter(df_os_risk, SubjectID %in% int_id) %>% distinct() %>%
  mutate(Res = case_when(Res == 0 ~ "non-responder",
                         Res == 1 ~ "responder"))

# Get numeber of Res and Non-res
xx <- df_os_risk %>% filter(Visit == "V1")
table(xx$Res)

p_os <- 
  ggplot(df_os_risk, aes(x=Visit, y=Sum_RS)) +
  geom_point(size=2) +
  geom_line(aes(group=SubjectID)) +
  facet_wrap(~Res, ncol=2) +
  geom_signif(
    comparisons = list(c("V1", "V2")),
    map_signif_level = FALSE, textsize = 6,
    test = "t.test",
    test.args = list(paired = TRUE),
  ) + 
  theme_bw() +
  theme(legend.position = "none",
          panel.grid.major = element_line(linetype = "blank"), 
          panel.grid.minor = element_line(linetype = "blank")) +
  NULL

p_os

ggsave(p_os, file=file.path(wd$output, "RS_OS_response.pdf"), device = "pdf", width = 12, height = 8)
```

Make plot nicer

```{r}
# https://yjunechoe.github.io/posts/2020-07-13-geom-paired-raincloud/
p_OS2 <- 
df_os_risk %>% 
  arrange(Visit) %>%
  ggplot(aes(x = Visit, y = Sum_RS, fill = Visit)) +
  #geom_paired_raincloud(alpha = .5, trim=TRUE, scale="count") +
  geom_point(aes(group = SubjectID),
             #position = position_nudge(c(.15, -.15)),
             alpha = 1) +
  geom_line(aes(group = SubjectID),
            #position = position_nudge(c(.13, -.13)),
            linetype = 3) +
  geom_boxplot(position = position_nudge(c(-.08, .08)),
               alpha = .5, width = .05, outlier.shape = " ") +
  facet_wrap(~Res) +
  theme_bw() +
  theme(legend.position = "none",
        panel.grid.major = element_line(linetype = "blank"), 
        panel.grid.minor = element_line(linetype = "blank")) +
  geom_signif(
    comparisons = list(c("V1", "V2")),
    map_signif_level = TRUE, textsize = 6,
    test = "t.test",
    test.args = list(paired = TRUE),
  ) + 
  NULL

ggsave(p_OS2, file=file.path(wd$output, "RS_OS_response_2.pdf"), device = "pdf", width = 12, height = 8)
```


For PFS


```{r}
df_pfs_risk <- filter(df_pfs_risk, SubjectID %in% int_id) %>% distinct() %>%
  mutate(Res = case_when(Res == 0 ~ "non-responder",
                         Res == 1 ~ "responder"))

p_pfs <-
  ggplot(df_pfs_risk, aes(x=Visit, y=Sum_RS)) +
  geom_point(size=2) +
  geom_line(aes(group=SubjectID)) +
  facet_wrap(~Res, ncol=2) +
  geom_signif(
    comparisons = list(c("V1", "V2")),
    map_signif_level = FALSE, textsize = 6,
    test = "t.test",
    test.args = list(paired = TRUE),
  ) + 
  theme_bw() +
  theme(legend.position = "none",
          panel.grid.major = element_line(linetype = "blank"), 
          panel.grid.minor = element_line(linetype = "blank")) +
  NULL
p_pfs

ggsave(p_pfs, file=file.path(wd$output, "RS_PFS_response.pdf"), device = "pdf", width = 12, height = 8)
```

Make plot nicer

```{r}
# https://yjunechoe.github.io/posts/2020-07-13-geom-paired-raincloud/
p_PFS2 <- 
df_pfs_risk %>% 
  arrange(Visit) %>%
  ggplot(aes(x = Visit, y = Sum_RS, fill = Visit)) +
  #geom_paired_raincloud(alpha = .5, trim=TRUE, scale="count") +
  geom_point(aes(group = SubjectID),
             #position = position_nudge(c(.15, -.15)),
             alpha = 1) +
  geom_line(aes(group = SubjectID),
            #position = position_nudge(c(.13, -.13)),
            linetype = 3) +
  geom_boxplot(position = position_nudge(c(-.08, .08)),
               alpha = .5, width = .05, outlier.shape = " ") +
  facet_wrap(~Res) +
  theme_bw() +
  theme(legend.position = "none",
        panel.grid.major = element_line(linetype = "blank"), 
        panel.grid.minor = element_line(linetype = "blank")) +
  geom_signif(
    comparisons = list(c("V1", "V2")),
    map_signif_level = FALSE, textsize = 6,
    test = "t.test",
    test.args = list(paired = TRUE),
  ) + 
  NULL

ggsave(p_PFS2, file=file.path(wd$output, "RS_PFS_response_2.pdf"), device = "pdf", width = 12, height = 8)
```


Save the risk score information


```{r}
# OS
t1 <- df_risk_v1_OS_solid %>% mutate(Visit = "V1") %>% select(PatientID, Sum_RS, Visit, Res, RS_binr) %>% left_join(tmp_df_1) %>%
  rename(RS_OS = Sum_RS,
         RS_OS_bin = RS_binr) 
t2 <- df_risk_v1_PFS_solid %>% mutate(Visit = "V1") %>% select(PatientID, Sum_RS, Visit, Res, RS_binr) %>% left_join(tmp_df_1) %>%
  rename(RS_PFS = Sum_RS,
         RS_PFS_bin = RS_binr)  %>%
  select(PatientID, RS_PFS, RS_PFS_bin)

df_visit_1_rs <- left_join(t1, t2)

write.csv(df_visit_1_rs, file.path(wd$output, "Visit_1_RS.csv"))

# PFS
t1 <- df_risk_v2_OS_solid %>% mutate(Visit = "V2") %>% select(PatientID, Sum_RS, Visit, Res, RS_binr) %>% left_join(tmp_df_2) %>%
    rename(RS_OS = Sum_RS,
         RS_OS_bin = RS_binr) 
t2 <- df_risk_v2_PFS_solid %>% mutate(Visit = "V2") %>% select(PatientID, Sum_RS, Visit, Res, RS_binr) %>% left_join(tmp_df_2) %>%
  rename(RS_PFS = Sum_RS,
         RS_PFS_bin = RS_binr)  %>%
  select(PatientID, RS_PFS, RS_PFS_bin)

df_visit_2_rs <- left_join(t1, t2)


write.csv(df_visit_2_rs, file.path(wd$output, "Visit_2_RS.csv"))
```

### ctDNA compare

We compare ctDNA before and after treatment

```{r}
ct_info <- df_tumor_p3 %>% select(PatientID, ctDNA, ct_cut)
df_pfs_risk_ct <- df_pfs_risk %>% left_join(ct_info)

# Make ctDNA plot
p_ctDNA <- 
df_pfs_risk_ct %>% 
  arrange(Visit) %>%
  ggplot(aes(x = Visit, y = ctDNA, fill = Visit)) +
  #geom_paired_raincloud(alpha = .5, trim=TRUE, scale="count") +
  geom_point(aes(group = SubjectID),
             #position = position_nudge(c(.15, -.15)),
             alpha = 1) +
  geom_line(aes(group = SubjectID),
            #position = position_nudge(c(.13, -.13)),
            linetype = 3) +
  geom_boxplot(position = position_nudge(c(-.08, .08)),
               alpha = .5, width = .05, outlier.shape = " ") +
  facet_wrap(~Res) +
  theme_bw() +
  theme(legend.position = "none",
        panel.grid.major = element_line(linetype = "blank"), 
        panel.grid.minor = element_line(linetype = "blank")) +
  geom_signif(
    comparisons = list(c("V1", "V2")),
    map_signif_level = FALSE, textsize = 6,
    test = "t.test",
    test.args = list(paired = TRUE),
  ) + 
  NULL

ggsave(p_ctDNA, file=file.path(wd$output, "ctDNA_response_2.pdf"), device = "pdf", width = 12, height = 8)
```





# Public risk score

Lets calculate the risk score using an independent data. For this we use this [paper](https://www.pnas.org/doi/10.1073/pnas.1902651116) and the cBioPoral [link](https://www.cbioportal.org/study/summary?id=prad_su2c_2019). Its the stand up to cancer where we have treatment naive patients and the information on OS. 

Lets read clinical data and OS data into R

```{r}
# Where we manually select the pre-treatment
# We extracted the metadata from the supplemntary table of the original publication - https://www.ncbi.nlm.nih.gov/pubmed/31061129
d6_pub <- read_xlsx(file.path(wd$data, "public_data/pnas.1902651116.sd01_filtered.xlsx")) %>%
  # Filter to Adrenocarcioma samples and treatment Naive
  filter(`Pathology Classification` == "Adenocarcinoma" & 
           `Abiraterone (ABI) and Enzalutamide (ENZA) Exposure Status` == "Naive" &
           `Taxane exposure status` == "Naïve") 

# All 444 samples from cBioPortal 
d6_sub <- read_xlsx(file.path(wd$data, "public_data/cBioPortal_prostate/02_prad_su2c_2019_clinical_data.xlsx"))  %>%
  # Subset
  filter(`Patient ID` %in% d6_pub$`Subject ID`) %>%
  filter(`Abiraterone (ABI) and Enzalutamide (ENZA) Exposure Status` == "Naive") %>%
  # Add column to be used in cBioPortal
  mutate(Study_Sample_ID = paste(`Study ID`, `Sample ID`, sep= ":"))

writexl::write_xlsx(d6_sub, file.path(wd$output, "PNAS_2019_Naive.xlsx"))

```


Using the naive samples above, **query cBioPotal for the 11 genes alternations in search box**. Use the identifier from "Sample ID" column in the excel file above to query cBioPortal.  

cBioPortal provdes two ways of downlading the data. The first from the OncoPrint plot itself you can download the Tabluar data. The second from the "Download" tab under "Downloadable Data Files". 

An alternate way to download data from cBioPortal is without querying the genes of interest. On the main [page](https://www.cbioportal.org/study/summary?id=prad_su2c_2019), we can go to the download symbol button at the top of the page. The direct link to download all data from this study is - https://cbioportal-datahub.s3.amazonaws.com/prad_su2c_2019.tar.gz.





```{r}
# First file from OoncoPrint plot
# We cleaned the file to make amp:-1 and del :1
d1 <- read_xlsx(file.path(wd$data, "public_data/cBioPortal_prostate/SU2C_perPatient_CNV/PATIENT_DATA_oncoprint_edited.xlsx"))


# Read the 2nd file
d2 <- read.delim(file.path(wd$data, "public_data/cBioPortal_prostate/SU2C_perPatient_CNV/cna_edit.txt"))

# When we read in the full data of cBioportal
# d2 <- read.delim(file.path(wd$data, "public_data/cBioPortal_prostate/SU2C_perPatient_CNV/data_cna.txt")) %>% 
#   filter(Hugo_Symbol %in% gene_or)
# Need to use readr package, other wise colnames will have .
d2 <- read_delim(file.path(wd$data, "public_data/cBioPortal_prostate/SU2C_perPatient_CNV/data_cna.txt")) %>%
  filter(Hugo_Symbol %in% gene_or)
# Subset to the 126 treatment naive patients
pt_naive <- d6_sub$`Sample ID`

gene_name <- d2$Hugo_Symbol
d2 <- d2[,pt_naive]
d2 <- t(d2)
colnames(d2) <- gene_name

# The matching patient ID and smample id
d3 <- read.delim(file.path(wd$data, "public_data/cBioPortal_prostate/SU2C_perPatient_CNV/alterations_across_samples.tsv")) %>% 
  select(Sample_ID, Patient_ID)
  
```

Lets convert the call to binary, similar to the PROMOTE cohort

```{r}
d2 <- d2 %>% as.data.frame() %>% tibble::rownames_to_column(var = "Sample_ID")
# Exclude with no data on all genes
d2 <- d2 %>% na.omit()

# Rearrange the df to make it easier
d2 <- d2 %>% select(Sample_ID, one_of(gene_or))

d2 <- d2 %>%
  mutate(
    across(AR:NOTCH1, ~ case_when(
      . == 0 ~ 0,
      . == -1 ~ 0,
      . == -2 ~ 0,
      . == 1 ~ 1,
      . == 2 ~ 1)),
    across(TMPRSS2:'NKX3-1', ~ case_when(
      . == 0 ~ 0,
      . == -1 ~ -1,
      . == -2 ~ -1,
      . == 1 ~ 0,
      . == 2 ~ 0)))

```



The first data has OS information extracted from cBioPortal, we also have OS information from the publication. We see if these matches


```{r}
d6_sub2 <- d6_sub %>% 
  rename(Patient = 'Patient ID',
         Sample_ID = 'Sample ID',
         OS_supp = 'Overall Survival (Months)',
         OS_status_supp = 'Overall Survival Status',
         AR_Score = 'AR Score',
         NEPC_Score = 'NEPC Score') %>%
  dplyr::select(Sample_ID, OS_supp,OS_status_supp,AR_Score,NEPC_Score) %>%
  mutate(OS_status_supp = gsub(":.*", "", OS_status_supp))

# df_1 <- d1 %>%  left_join(d6_sub2) %>%  # Yes its the same 
#   mutate(across(AR:ZBTB16, ~ case_when(
#     . == -1 ~ 1,
#     TRUE ~ . ))) %>% 
#     mutate(OS_supp = as.numeric(OS_supp),
#          OS_status_supp = as.numeric(OS_status_supp)) %>%
#   filter(!is.na(OS_supp)) %>%
#   select(-OS, -Death) %>%
#   rename(Patient_ID = Patient)

```


Add OS to the 2nd file, but first add the patient ID

```{r}
df_2 <- d2 %>%  left_join(d3) %>%  
    mutate(across(AR:'NKX3-1', ~ case_when(
    . == -1 ~ 1,
    TRUE ~ . ))) %>% 
  left_join(d6_sub2, by = c("Sample_ID" = "Sample_ID")) %>%
  mutate(OS_supp = as.numeric(OS_supp),
         OS_status_supp = as.numeric(OS_status_supp)) %>%
  filter(!is.na(OS_supp))


```

Now lets calculate the risk score and run survival analysis. Modify the top to df_1 or df_2 based on the input data. 

```{r,message=FALSE, warning=FALSE}
# Calculate cox
in_df <- df_2
cox_results <- list()
gene_col <- in_df %>% select(AR:'NKX3-1' ) %>% colnames()
for (gene in colnames(in_df)[2:12]) {
  
  tmp_df <- in_df %>% select(!!as.name(gene), OS_supp, OS_status_supp) %>% # ------- CAHNGE TO OS = Death OR PFS - Prog
    rename(gene = !!as.name(gene))
  cox_model <- summary(coxph(Surv(OS_supp, OS_status_supp) ~ gene, data = tmp_df))
  
  cox_results[[gene]] <- cox_model$coefficients
}

cox_result_v1 <- do.call(rbind, cox_results) %>% as.data.frame()
rownames(cox_result_v1) <- gene_col

df_cox_v1 <- tibble(rownames_to_column(cox_result_v1, var = "Gene")) %>%
  rename(pVal = 'Pr(>|z|)') %>%
  filter(!is.na(coef))

# Cacluate Risk score
# All patient
pat_all <- in_df$Patient_ID
pat_sum_list <- list()

for (i in seq_along(pat_all)){
  patID <- pat_all[i]
  pat_1 <- in_df %>% filter(Patient_ID == patID) %>% select(Patient_ID, df_cox_v1$Gene) %>% 
    pivot_longer(!Patient_ID, names_to = "Gene", values_to = "CNR") %>%
    left_join(df_cox_v1) %>%
    # Calculate risk score
    mutate(RS = coef * CNR)
  # get the sum
  pat_sum_list[[i]] <- sum(pat_1$RS)
}

df_pat_sum <- data.frame(
  Patient_ID = pat_all,
  Sum_RS = unlist(pat_sum_list)
)


df_pat_sum <- df_pat_sum %>% 
  mutate(RS_binr = case_when(Sum_RS > median(Sum_RS) ~ "high",
                             Sum_RS <= median(Sum_RS) ~ "low"))

# Calculate and plot OS
df_risk_public <- in_df %>% left_join(df_pat_sum)

write.csv(df_risk_public, file.path(wd$output, "cBioPortal_riskScore.csv"))

# Fit survival data using the Kaplan-Meier method
surv_object <- Surv(time = df_risk_public$OS_supp, event = df_risk_public$OS_status_supp) # ------- CAHNGE TO OS = Death OR PFS - Prog
# Fit Kaplan-Meier curve
fit1 <- survfit(surv_object ~ RS_binr, data = df_risk_public)
p_OS <- ggsurvplot(fit1, data = df_risk_public, pval = TRUE, risk.table=TRUE)
p_OS

```

Make figure nicer

```{r}
pdf(file.path(wd$output, "OS_cBioPortal.pdf"), width = 12, height = 8)
ggsurvplot(fit1, data = df_risk_public, pval = TRUE, risk.table=TRUE,
           conf.int = TRUE,
           palette = c("#e41a1c", "#7F7F7F"))
dev.off()
```


## Other risk score

In the cBioportal dataset, we had two other clinical information, the NEPC and AR score. 

We binarize this score based on median

```{r}
df_risk_public_AR <- df_risk_public %>% 
  filter(!is.na(AR_Score)) %>% 
  filter(!AR_Score == "NA") %>% 
  mutate(AR_Score = as.numeric(AR_Score),
         AR_Score_bin = case_when(
           AR_Score > median(AR_Score) ~ "high",
           AR_Score <= median(AR_Score) ~ "low"))

  
surv_object <- Surv(time = df_risk_public_AR$OS_supp, event = df_risk_public_AR$OS_status_supp)    
```


```{r}
# Fit survival data using the Kaplan-Meier method
surv_object <- Surv(time = df_risk_public_AR$OS_supp, event = df_risk_public_AR$OS_status_supp) 
# Fit Kaplan-Meier curve
fit1 <- survfit(surv_object ~ AR_Score_bin, data = df_risk_public_AR)
p_OS_AR <- ggsurvplot(fit1, data = df_risk_public_AR, pval = TRUE, risk.table=TRUE, conf.int = TRUE)
p_OS_AR
```


Repeat for NEPC

```{r}
df_risk_public_NEPC <- df_risk_public %>% 
  filter(!is.na(NEPC_Score)) %>% 
  filter(!NEPC_Score == "NA") %>% 
  mutate(NEPC_Score = as.numeric(NEPC_Score),
         NEPC_Score_bin = case_when(
           NEPC_Score > median(NEPC_Score) ~ "high",
           NEPC_Score <= median(NEPC_Score) ~ "low"))

  
surv_object <- Surv(time = df_risk_public_NEPC$OS_supp, event = df_risk_public_NEPC$OS_status_supp)   

# Fit survival data using the Kaplan-Meier method
surv_object <- Surv(time = df_risk_public_NEPC$OS_supp, event = df_risk_public_NEPC$OS_status_supp) 
# Fit Kaplan-Meier curve
fit1 <- survfit(surv_object ~ NEPC_Score_bin, data = df_risk_public_NEPC)
p_OS_NEPC <- ggsurvplot(fit1, data = df_risk_public_NEPC, pval = TRUE, risk.table=TRUE, conf.int = TRUE)
p_OS_NEPC
```

Save the plot

```{r}
pdf(file.path(wd$output, "OS_cBioPortal_AR_NEPC.pdf"), width = 12, height = 8)
p_OS_AR
p_OS_NEPC
dev.off()
```

Save the risk score information, alongside the AR and NEPC score

```{r}
t1 <- df_risk_public_NEPC %>% select(Sample_ID, NEPC_Score, NEPC_Score_bin)
t2 <- df_risk_public_AR %>% select(Sample_ID, AR_Score, AR_Score_bin)

df_risk_public_full <- df_risk_public %>% 
  select(-AR_Score, -NEPC_Score) %>% 
  left_join(t1) %>% 
  left_join(t2)


write.csv(df_risk_public_full, file.path(wd$output, "cBioPortal_riskScore.csv"))

```


## Pre-calculated weights

What if we used the pre-calculated weights

```{r, warning=FALSE, message=FALSE}
# We stotred the visit 1 coeeficients - df_cox_visit_1
in_df <- df_2


pat_all <- in_df$Patient_ID
pat_sum_list <- list()

for (i in seq_along(pat_all)){
  patID <- pat_all[i]
  pat_1 <- in_df %>% filter(Patient_ID == patID) %>% select(Patient_ID, df_cox_v1$Gene) %>% 
    pivot_longer(!Patient_ID, names_to = "Gene", values_to = "CNR") %>%
    left_join(df_cox_visit_1) %>%
    # Calculate risk score
    mutate(RS = coef * CNR)
  # get the sum
  pat_sum_list[[i]] <- sum(pat_1$RS)
}

df_pat_sum <- data.frame(
  Patient_ID = pat_all,
  Sum_RS = unlist(pat_sum_list)
)



df_pat_sum <- df_pat_sum %>% 
  mutate(RS_binr = case_when(Sum_RS > median(Sum_RS) ~ "high",
                             Sum_RS <= median(Sum_RS) ~ "low"))

# Calculate and plot OS
df_risk_public <- in_df %>% left_join(df_pat_sum)

# Fit survival data using the Kaplan-Meier method
surv_object <- Surv(time = df_risk_public$OS_supp, event = df_risk_public$OS_status_supp) # ------- CAHNGE TO OS = Death OR PFS - Prog
# Fit Kaplan-Meier curve
fit1 <- survfit(surv_object ~ RS_binr, data = df_risk_public)
p_OS <- ggsurvplot(fit1, data = df_risk_public, pval = TRUE, risk.table=TRUE)
p_OS
```


# Wyatt risk score


Manish reached out to Wyatt for his patients clinical outcome and copy number scores. We read in the data provided. 

```{r}
dat_wyat <- readxl::read_xlsx(file.path(wd$data, "validation_cohort_kohli_feb_14_2024_edit.xlsx"))
```

Make the copy number calls on selected genes. Wyatt have data only from 8 of the 11 genes we are interested in.

```{r}
# We use 0.3 / -0.3 threshold
thresh_val <- 0.3

# Note that some of the samples does not have CNV calls for TMPRSS2. 
# So we will keep a data frame with all samples and a second data frame excluding these samples

# Full samples
dat_wyatt_full <- dat_wyat %>% 
  rename(OS = 'OS (days) measured from initiation of 1L mCRPC treatment',
         Death = 'Survival status (alive or lost to follow-up = 0; dead = 1)',
         PFS = 'PSA PFS (days) measured from initiation of 1L mCRPC treatment',
         Prog = 'PSA progression (no = 0; yes = 1)',
         Txt = '1L mCRPC systemic treatment (abiraterone = 1; enzalutamide = 2; docetaxel = 3; cabazitaxel = 4; other = 5; radium-223 = 6)',
         ALP = 'ALP per ULN',
         LDH = 'LDH per ULN',
         Hb = 'Hb (g/L)') %>% 
  select(Patient_ID, 
         AR, MYC, PIK3CA, PIK3CB, 
         NKX3.1, TMPRSS2, TP53, ZBTB16,
         OS, Death, PFS, Prog,
         Txt, ALP, LDH, Hb) %>% 
  mutate(
    OS = round(OS/30.417, digit=2),
    PFS = round(PFS/30.417, digit=2),
    across(AR:PIK3CB, ~ case_when(
      . >= thresh_val ~ 1,
      TRUE ~ 0)),
    across(NKX3.1:ZBTB16, ~ case_when(
      . <= -thresh_val ~ -1,
      is.na(.) ~ NA,
      TRUE ~ 0))
    ) %>% as.data.frame()

dat_wyatt <- dat_wyatt_full %>% 
  # Note that some of the samples does not have CNV calls for TMPRSS2. So we will exclude these samples
  filter(!is.na(TMPRSS2))


# # Filter to those that was treated with Abi
# #dat_wyatt_abi <- dat_wyatt %>% filter(Txt == 1)
# dat_wyatt_abi <- dat_wyatt 
```



## Risk score

### OS

Calculate risk score based on OS

```{r,warning=FALSE, message=FALSE}
# Calculate cox
in_df <- dat_wyatt
cox_results <- list()
gene_col <- in_df %>% select(AR:ZBTB16 ) %>% colnames()
for (gene in colnames(in_df)[2:9]) {
  
  tmp_df <- in_df %>% select(!!as.name(gene), OS, Death) %>% # ------- CAHNGE TO OS = Death OR PFS - Prog
    rename(gene = !!as.name(gene))
  cox_model <- summary(coxph(Surv(OS, Death) ~ gene, data = tmp_df))
  
  cox_results[[gene]] <- cox_model$coefficients
}

cox_result_v1 <- do.call(rbind, cox_results) %>% as.data.frame()
rownames(cox_result_v1) <- gene_col

df_cox_v1 <- tibble(rownames_to_column(cox_result_v1, var = "Gene")) %>%
  rename(pVal = 'Pr(>|z|)') %>%
  filter(!is.na(coef))

# Cacluate Risk score
# All patient
pat_all <- in_df$Patient_ID
pat_sum_list <- list()

for (i in seq_along(pat_all)){
  patID <- pat_all[i]
  pat_1 <- in_df %>% filter(Patient_ID == patID) %>% select(Patient_ID, df_cox_v1$Gene) %>% 
    pivot_longer(!Patient_ID, names_to = "Gene", values_to = "CNR") %>%
    left_join(df_cox_v1) %>%
    # Calculate risk score
    mutate(RS = coef * CNR)
  # get the sum
  pat_sum_list[[i]] <- sum(pat_1$RS)
}

df_pat_sum <- data.frame(
  Patient_ID = pat_all,
  Sum_RS = unlist(pat_sum_list)
)


df_pat_sum <- df_pat_sum %>% 
  mutate(RS_binr = case_when(Sum_RS > median(Sum_RS) ~ "high",
                             Sum_RS <= median(Sum_RS) ~ "low"))




# Calculate and plot OS
df_risk_public <- in_df %>% left_join(df_pat_sum)

# Save csv
write.csv(df_risk_public, file.path(wd$output, "Wyatt_riskScore.csv"))


df_risk_os_wyatt <- df_risk_public
# Fit survival data using the Kaplan-Meier method
surv_object <- Surv(time = df_risk_public$OS, event = df_risk_public$Death) # ------- CAHNGE TO OS = Death OR PFS - Prog
# Fit Kaplan-Meier curve
fit1 <- survfit(surv_object ~ RS_binr, data = df_risk_public)
p_OS <- ggsurvplot(fit1, data = df_risk_public, pval = TRUE, risk.table=TRUE,
                    conf.int = TRUE,
                    palette = c("#e41a1c", "#7F7F7F"))
p_OS


pdf(file.path(wd$output, "OS_WyattData_all.pdf"), width = 12, height = 8)
p_OS
dev.off()
```
### PFS 

Now we want to calculate the risk score based on PFS. We sepecifically want to look at the PFS for patients treated with 1) abiraterone and 2) Enzalutamide.

First calculate a global PFS risk score of all patients.

```{r,warning=FALSE, message=FALSE}
# Calculate cox
in_df <- dat_wyatt
cox_results <- list()
gene_col <- in_df %>% select(AR:ZBTB16 ) %>% colnames()
for (gene in colnames(in_df)[2:9]) {
  
  tmp_df <- in_df %>% select(!!as.name(gene), PFS, Prog) %>% # ------- CAHNGE TO OS = Death OR PFS - Prog
    rename(gene = !!as.name(gene))
  cox_model <- summary(coxph(Surv(PFS, Prog) ~ gene, data = tmp_df))
  
  cox_results[[gene]] <- cox_model$coefficients
}

cox_result_v1 <- do.call(rbind, cox_results) %>% as.data.frame()
rownames(cox_result_v1) <- gene_col

df_cox_v1 <- tibble(rownames_to_column(cox_result_v1, var = "Gene")) %>%
  rename(pVal = 'Pr(>|z|)') %>%
  filter(!is.na(coef))

# Cacluate Risk score
# All patient
pat_all <- in_df$Patient_ID
pat_sum_list <- list()

for (i in seq_along(pat_all)){
  patID <- pat_all[i]
  pat_1 <- in_df %>% filter(Patient_ID == patID) %>% select(Patient_ID, df_cox_v1$Gene) %>% 
    pivot_longer(!Patient_ID, names_to = "Gene", values_to = "CNR") %>%
    left_join(df_cox_v1) %>%
    # Calculate risk score
    mutate(RS = coef * CNR)
  # get the sum
  pat_sum_list[[i]] <- sum(pat_1$RS)
}

df_pat_sum <- data.frame(
  Patient_ID = pat_all,
  Sum_RS = unlist(pat_sum_list)
)


df_pat_sum <- df_pat_sum %>% 
  mutate(RS_binr = case_when(Sum_RS > median(Sum_RS) ~ "high",
                             Sum_RS <= median(Sum_RS) ~ "low"))

# Calculate and plot OS
df_risk_public <- in_df %>% left_join(df_pat_sum)
df_risk_pfs_wyatt <- df_risk_public

# Save csv
write.csv(df_risk_public, file.path(wd$output, "Wyatt_riskScore_PFS.csv"))

# Fit survival data using the Kaplan-Meier method
surv_object <- Surv(time = df_risk_public$PFS, event = df_risk_public$Prog) # ------- CAHNGE TO OS = Death OR PFS - Prog
# Fit Kaplan-Meier curve
fit1 <- survfit(surv_object ~ RS_binr, data = df_risk_public)
p_PFS <- ggsurvplot(fit1, data = df_risk_public, pval = TRUE, risk.table=TRUE,
                    conf.int = TRUE,
                    palette = c("#e41a1c", "#7F7F7F"))
p_PFS


pdf(file.path(wd$output, "PFS_WyattData.pdf"), width = 12, height = 8)
p_PFS
dev.off()
```

Save the risk score values (OS & PFS) into a table.

```{r}
os_sub <- df_risk_os_wyatt %>% 
  select(Patient_ID, Sum_RS, RS_binr) %>% 
  dplyr::rename(OS_RS = Sum_RS,
                OS_bin = RS_binr)

pfs_sub <- df_risk_pfs_wyatt %>% 
  select(Patient_ID, Sum_RS, RS_binr) %>% 
  dplyr::rename(PFS_RS = Sum_RS,
                PFS_bin = RS_binr)

wyatt_rs <- dat_wyatt_full %>% left_join(os_sub) %>% left_join(pfs_sub)

write.csv(wyatt_rs, file.path(wd$output, "Wyatt_MetaData.csv"))
```



#### Abi

To obtain abi specific plot, we subset the risk score and PFS for only abi treated patients

```{r}
# Data frame of only abi
df_risk_public_abi <- df_risk_public %>% 
  filter(Txt == 1)

# Fit survival data using the Kaplan-Meier method
surv_object <- Surv(time = df_risk_public_abi$PFS, event = df_risk_public_abi$Prog) # ------- CAHNGE TO OS = Death OR PFS - Prog
# Fit Kaplan-Meier curve
fit_abi <- survfit(surv_object ~ RS_binr, data = df_risk_public_abi)
p_PFS_abi <- ggsurvplot(fit_abi, data = df_risk_public_abi, pval = TRUE, risk.table=TRUE,
                        conf.int = TRUE,
                        palette = c("#e41a1c", "#7F7F7F"))
p_PFS_abi


pdf(file.path(wd$output, "PFS_WyattData_abi.pdf"), width = 12, height = 8)
p_PFS_abi
dev.off()
```
#### Enza

Repeat for enzalutamide

```{r}
# Data frame of only enza
df_risk_public_enz <- df_risk_public %>% 
  filter(Txt == 2)

# Fit survival data using the Kaplan-Meier method
surv_object <- Surv(time = df_risk_public_enz$PFS, event = df_risk_public_enz$Prog) # ------- CAHNGE TO OS = Death OR PFS - Prog
# Fit Kaplan-Meier curve
fit_enz <- survfit(surv_object ~ RS_binr, data = df_risk_public_enz)
p_PFS_enz <- ggsurvplot(fit_enz, data = df_risk_public_enz, pval = TRUE, risk.table=TRUE,
                        conf.int = TRUE,
                        palette = c("#e41a1c", "#7F7F7F"))
p_PFS_enz


pdf(file.path(wd$output, "PFS_WyattData_enz.pdf"), width = 12, height = 8)
p_PFS_enz
dev.off()
```

## Other factors

Lets try to do the cox model on other factors, not only genes

Fist run uni-variate analysis

```{r}
df_risk_public2 <- df_risk_os_wyatt %>% 
  mutate(RS_binr = factor(RS_binr, levels=c("low", "high")))

df_in <- df_risk_public2
df_in_low <- df_in %>% filter(RS_binr == "low")
df_in_high <- df_in %>% filter(RS_binr == "high")


surv_object <- Surv(time = df_in$OS, event = df_in$Death) # ------- CHANGE TO OS = Death OR PFS - Prog
# --------- #
# Uni-variate
# --------- #
# # Fit a Cox proportional hazards model
# fit.coxph <- coxph(surv_object ~ RS_binr,
#                    data = df_in)
# # Extact values
# hr <- exp(coef(fit.coxph))
# hr
# ggforest(fit.coxph, data = df_in)
# 
# 
# xx <- survfit2(Surv(OS, Death) ~ RS_binr + ALP, data = df_in)
# 
# # Plot
# ggsurvfit(xx, linewidth = 1)
# 
#  #Plot survival curves for each variable
#  ggsurvplot(survfit(fit.coxph), data = df_in_low, title = "Univariate Survival Analysis",
#             pval = TRUE, risk.table=TRUE)

# --------- #
# Uni-variate
# --------- #

# Fit a Cox proportional hazards model
fit.coxph2.1 <- coxph(surv_object ~  ALP,
                   data = df_in)
fit.coxph2.2 <- coxph(surv_object ~  LDH,
                   data = df_in)
fit.coxph2.3 <- coxph(surv_object ~  Hb,
                   data = df_in)
fit.coxph2.4 <- coxph(surv_object ~  RS_binr,
                   data = df_in)
# hr2 <- exp(coef(fit.coxph2))
# ggforest(fit.coxph2, data = df_in)

pdf(file.path(wd$output, "HR_OS_Wyatt_Uni_RS.pdf"), width = 12, height = 8)
ggforest(fit.coxph2.1, data = df_in)
ggforest(fit.coxph2.2, data = df_in)
ggforest(fit.coxph2.3, data = df_in)
ggforest(fit.coxph2.4, data = df_in)
dev.off()


# Extract results
r1 <- summary(fit.coxph2.1)

# --------- #
# Multi-variate
# --------- #
fit.coxph3 <- coxph(surv_object ~ ALP + LDH + Hb + RS_binr,
                   data = df_in)
# hr3 <- exp(coef(fit.coxph3))
# ggforest(fit.coxph2, data = df_in)

pdf(file.path(wd$output, "HR_OS_Wyatt_Multi_RS.pdf"), width = 12, height = 8)
ggforest(fit.coxph3, data = df_in)
dev.off()
```


Extract resutls of individual

```{r}
extract_coxph_results <- function(cox_models) {
  result_list <- lapply(cox_models, function(model) {
    # Extract coefficients and confidence intervals
    coef_summary <- summary(model)$coefficients
    coef_summary2 <- summary(model)$conf.int
    coef_names <- row.names(coef_summary)
    coef <- coef_summary[, 1]
    lower_bound <- coef_summary2[, 3]
    upper_bound <- coef_summary2[, 4]
    p_value <- coef_summary[, 5]
    
    # Exponentiate coefficients for hazard ratios (HRs)
    exp_coef <- exp(coef)
    
    # Create a data frame
    result_df <- data.frame(
      Coefficient = coef,
      Lower_Bound = lower_bound,
      Upper_Bound = upper_bound,
      P_Value = p_value,
      HR = exp_coef
    )
    row.names(result_df) <- coef_names
    return(result_df)
  })
  
  # Combine results into a single data frame
  combined_results <- do.call(rbind, result_list)
  
  return(combined_results)
}

# List of Cox models
cox_models <- list(fit.coxph2.1, fit.coxph2.2, fit.coxph2.3, fit.coxph2.4)

# Extract results for the multiple univariate Cox models 
results_uni <- extract_coxph_results(cox_models) %>% 
  tibble::rownames_to_column(var = "Features")
# Save as csv
writexl::write_xlsx(results_uni, file.path(wd$output, "UniVariate_cox.xlsx"))


# Extract results for multi-variate Cox model
coef_summary <- summary(fit.coxph3)$coefficients
results_multi <- data.frame(
  Coefficient = coef_summary[, 1],
  Lower_Bound = summary(fit.coxph3)$conf.int[, 3],
  Upper_Bound =summary(fit.coxph3)$conf.int[, 4],
  P_Value = coef_summary[, 5],
  HR =coef_summary[, 2]
) %>% 
  tibble::rownames_to_column(var = "Features")

writexl::write_xlsx(results_multi, file.path(wd$output, "MultiVariate_cox.xlsx"))
```




## Responders

What if we do the risk score difference between responder and non responders

```{r}

```

