---
title: "Parsing XML"
date: "2022-11-28"
output: 
  html_document:
    toc: true
    toc_float:
      collapsed: true
    toc_depth: 3
    number_sections: true
    theme: lumen
---

# Background

We want to mine clinicaltrials.gov[https://clinicaltrials.gov/] and extract serious adverse events after adminstration of immunotherapy drugs. The entire set of clinical trials data can be downloaded at https://clinicaltrials.gov/ct2/resources/download. Navigate to the section *Downloading All Study Record Content for Analysis*. 

# Objectives

1. Familiarize with XML package
    + Analyse 1 XML file 
2. Apply the workflow to all XML file downloaded
3. Save the DF
  

# Pre-processing 

## Loading packages

```{r, message=FALSE, warning=FALSE}
library(here)
library(tidyverse)
library(XML)
library(xml2)
library(methods) #install.packages("methods")
library(stringr)
```


## Directories

```{r}
wd <- list()
wd$main <- here()
wd$data <- file.path(wd$main, "data")
wd$dataP <- file.path(wd$data, "processed")
wd$output <- file.path(wd$main, "output")
wd$script <- file.path(wd$main, "scripts")
```


Create directories

```{r}
if (!file.exists(wd$output )) {
  dir.create(wd$output )
} else {
  print("Directory already exists")
}
```

Load functions

```{r}
source(file.path(wd$script, "functions.R"))
```

Drugs we are interested to query are listed in `functions.R`


# Objective 1

Here we try to read only 1 XML file. Mainly to set-up the scripts and see if we can extract the required info


```{r}
# Some example files
f1 <- file.path(wd$data, "clin_trials/AllPublicXML/NCT0383xxxx/NCT03834623.xml") # Example with results column (used to set up script)
#f1 <- file.path(wd$data, "clin_trials/AllPublicXML/NCT0002xxxx/NCT00025181.xml") # No results gives error in function, use as exaple to de-bug
#f1 <- file.path(wd$data, "clin_trials/AllPublicXML/NCT0230xxxx/NCT02301039.xml")
#f1 <- file.path(wd$data, "clin_trials/AllPublicXML/NCT0006xxxx/NCT00060372.xml") # This study have no results column
#f1 <- file.path(wd$data, "clin_trials/AllPublicXML/NCT0135xxxx/NCT01354431.xml") # This study has 3 arms
#f1 <- file.path(wd$data, "clin_trials/AllPublicXML/NCT0158xxxx/NCT01585987.xml") #  Two cohort but 3 serious events?
#f1 <- file.path(wd$data, "clin_trials/AllPublicXML/NCT0147xxxx/NCT01472081.xml") #  Study with 5 arms

# --- Some uncessfull or sub=optimal attempt --- $
#  those as the "data nodes", ignoring the others.
# f = system.file("exampleData", "tides.xml", package = "XML")
# doc = xmlParse(f1)
#d1 <- xmlToDataFrame(nodes = xmlChildren(xmlRoot(doc)[["clinical_results"]]))

#xml_data <- xmlToList(doc)

#location <- as.list(xml_data[["arm_group"]][["sponsors"]])
#rr <- as.list(xml_data[["clinical_results"]])

# Try to read nested xml [https://medium.com/geekculture/reading-xml-files-in-r-3122c3a2a8d9]

# study_data <- read_xml(f1) %>% as_list()
# study_data_sub <- study_data$clinical_study$clinical_results$reported_events
# study_tb = tibble::as_tibble(study_data) %>% tidyr::unnest_longer('clinical_study')
```


We found  an approach that is currently working. The script below extacts some information from the XML, but does not include the number of serious adverse event. 

```{r}
# Read the xml file
df <- read_xml(f1)

# ------------------ #
# Extracting study ID (clinical trials gov study ID)
# ------------------ #
xml_study_id <- 
  xml_text(
    xml_find_all(df, 
                 xpath = "//nct_id"))

# ------------------ #
# Extracting study status (has the study been completed)
# ------------------ #
xml_study_status <- 
  xml_text(
    xml_find_all(df, 
                 xpath = "//overall_status"))

# ------------------ #
# Extracting if the study reported the results (some studies don't report the results)
# ------------------ #
xml_result_status <- 
  xml_text(
    xml_find_all(df, 
                 xpath = "//clinical_results"))


# ------------------ #
# Extracting condition (cancer type)
# ------------------ #
xml_cond <- 
  xml_text(
    xml_find_all(df, 
                 xpath = "//condition"))

# ------------------ #
# Extracting drug given in each arm of the study (intervention type)
# ------------------ #
xml_intervention <- 
  xml_text(
    xml_find_all(df, 
                 xpath = "//intervention_name"))

# ------------------ #
# Extracting the study arm (how many variations to the study, different dose / drugs)
# ------------------ #
# Number of arm
xml_arm <- 
  xml_text(
    xml_find_all(df, 
                 xpath = "//number_of_arms"))

# Title / name of arm
xml_arm_name <- 
  xml_text(
    xml_find_all(df, 
                 xpath = "//arm_group/arm_group_label"))

# Name of group based on clinical results
xml_clin_group <- 
  xml_text(
    xml_find_all(df, 
                 xpath = "//clinical_results/participant_flow/group_list/group/title"))


# ------------------ #
# Extracting titles (irAE category)
# ------------------ #
# Extract the Main text
ser_event_title <- 
  xml_text(
    xml_find_all(df, 
                 xpath = "//reported_events/serious_events/category_list/category/title"))

# Extract the Sub title text
ser_event_subTitle <- 
  xml_text(
    xml_find_all(df, 
                 xpath = "//reported_events/serious_events/category_list/category/event_list/event/sub_title"))


```


Summarize the data as a data frame

```{r}
# Df 1 - Study info 
df_study_info <- data.frame(
  study_id = xml_study_id,
  study_status = xml_study_status,
  intervention = paste(xml_intervention, collapse = "_"),  
  condition = xml_cond,
  arm_num = xml_arm,
  arm_title = xml_arm_name,
  clin_group = xml_clin_group)
```



The next part of quering the XML file is to extract the number of serious adverse event. This is a bit more tricky to extract

```{r}
# ------------------ #
# Extracting number of cases with severe irAE 
# ------------------ #

# ------- DID NOT WORK ---------- #

# Approach 1
## Building a manual list and extracting info
#study_data <- read_xml(f1) %>% as_list()
#£study_data_sub <- study_data$clinical_study$clinical_results$reported_events$serious_events$category_list # Serious events columns

# If we take one serious event as an example
#tm_list_2 <- study_data_sub[1] # subset to one category 
#tm_list_3 <- tm_list_2$category$event_list$event # This is where the number is

# https://community.rstudio.com/t/extracting-from-nested-list/17016/2
#extract_col <- function(l, col) {
#  unlist(attributes(l)[col])
#}

#tm_list_3 %>% map_df(extract_col, "group_id") # Gives the group info
#tm_list_3 %>% map_df(extract_col, "subjects_affected") # Number of patients with the serious effects
#tm_list_3 %>% map_df(extract_col, "subjects_at_risk") # Number of patients in the study

# Approach 2 [NOT WORKING]
## Using a xml_find function to extract interested column
#ser_event_list <- 
#  xml_find_all(df, xpath = "//reported_events/serious_events/category_list/category") %>%
#  as_list() 


# Script not working below, we cant extract the interested attributes. So stick with approach one
#extract_col <- function(l, col) {
#  unlist(attributes(l$event_list$event$counts)[col])
#}
#ser_event_list[1] %>% map_df(extract_col, "group_id")

# ------- DID NOT WORK ---------- #


# ----- Current workflow ------ #

# Find the path which has the irAE event numbers
dat_path <- df %>% xml_find_all( '//counts') %>% xml_path # 42 lines
# Limit to serious effects
dat_path <- dat_path[grep("serious_events", dat_path) ]

#length(dat_path)
#head(dat_path)

# Now for one of the path found, extract the attributes
ser_event_num <- 
  xml_attrs(
    xml_find_all(df, 
                 xpath = "//reported_events/serious_events/category_list/category[12]/event_list/event/counts[1]"))
# ser_event_num # the output is a list of values

# Un-list the results
un_event <- unlist(ser_event_num)
group_id <- un_event["group_id"]
num_case <- un_event["subjects_affected"]
tot_case <- un_event["subjects_at_risk"]


# Now for all of the path found, extract the attributes using a loop

group_id <- vector()
num_case <- vector()
tot_case <- vector()


for (i in seq_along(dat_path)){
  ser_event_num <- 
    xml_attrs(xml_find_all(df, xpath = dat_path[i]))
  
  un_event <- unlist(ser_event_num)
  
  group_id[i] <- un_event["group_id"]
  num_case[i] <- un_event["subjects_affected"]
  tot_case[i] <- un_event["subjects_at_risk"]

}

df_out <- data.frame(
  group_id = group_id,
  num_case = num_case,
  tot_case = tot_case)


# Some studies list fewer arm than results (eg - NCT01585987.xml), in cases like this we get the max arm as defined by the study 
max_arm <- length(xml_arm_name)
# As arm is listed as E1, E2 E3 ....etc, we generate Ex according to max arm number
e_arm <- paste0("E", 1:max_arm)
# Filter event to only these category
df_out <- df_out %>%
  dplyr::filter(group_id %in% e_arm)



# ----- Current workflow ------ #
```


Once we have the  1) Study information 2) Number of serious adverse event,  we compile it into data frame

```{r}
# Df 1 - Study info 


# Df 2- serious event numbers
inf_fac <- max_arm # Some studies have multiple arms, resulting in each irAE listed multiple times. Find the number of arms

df_ser_number <- df_out %>%
  dplyr::mutate(study_id = xml_study_id,
                arm_title = rep(xml_arm_name, times=length(ser_event_subTitle)),
                serEvent = rep(ser_event_subTitle, each=inf_fac))
  
```

# Objective 2

Lets try to run this all in all studies in a folder. 

```{r}
list_files <- 
  #list.files(file.path(wd$data, "NCT_test/"),
  list.files(file.path(wd$data, "clin_trials/AllPublicXML/"),
  recursive = TRUE,
  pattern = ".xml",
  full.names = TRUE)
```


Some bash processing to reduce file number

Some studies have no result , so we use bash to identify which studies did not report their results. We  remove the study with no results. 

```{base, eval=FALSE}
cd /Users/zakiwilmot/Library/CloudStorage/Box-Box/Projects/ZW_005_irAE_database/analysis/data/clin_trials/AllPublicXML
grep("Results have not yet been posted", * > ../no_resutls.txt)
```

We also use bash to find any studies with these drugs

```{base, eval=FALSE}
# Takes soo long - Find a way to do this in R -- TO DO 
grep -E -i -r \
    "Ipilimumab|Yervoy|BMS-734016|MDX-010|MDX-101| \
    Nivolumab|Opdivo|ONO-4538|BMS-936558|MDX1106| \
    Pembrolizumab|Keytruda|MK-3475|lambrolizumab| \
    Atezolizumab|Tecentriq|MPDL3280A|RG7446| \
    Durvalumab|Imfinzi|MEDI4736|MEDI-4736| \
    Avelumab|Bavencio|MSB0010718C| \
    Cemiplimab|Libtayo|REGN-2810|REGN2810|cemiplimab-rwl| \
    Relatlimab|BMS-986016|relatlimab-rmbw| \
    Dostarlimab|Jemperli|TSR-042|WBP-285|dostarlimab-gxly| \
    Tremelimumab|Imjudo|tremelimumab-actl|ticilimumab|CP-675" \
    * > ../immuno_results.txt

```


Exclude processing files with no results. Read in the list of file with no results (based on our bash result above)

```{r}
no_res <- read.delim(file.path(wd$data, "processed/no_results.txt"), header = FALSE) %>%
  dplyr::mutate(fileName = gsub("\\.xml.*", "", V1),
                fileName = gsub(".*\\/NCT", "NCT", fileName))

# https://stackoverflow.com/questions/7597559/grep-using-a-character-vector-with-multiple-patterns
list_file_bn <- basename(list_files) 
list_file_bn <- gsub("\\.xml", "", list_file_bn)

# Df is easier

df_list <- data.frame(
  full_path = list_files,
  baseName = list_file_bn)

df_list_sub <- df_list %>% dplyr::filter(!baseName %in% no_res$fileName)
```

Read the result of files with the immuno drugs

```{r}
df_imu <- read.delim(file.path(wd$data, "processed/immuno_results.txt"), header = FALSE) %>%
  dplyr::mutate(fileName = gsub("\\.xml.*", "", V1),
                fileName = gsub(".*\\/NCT", "NCT", fileName))

imu <- unique(df_imu$fileName)

df_list_sub <- df_list_sub %>% dplyr::filter(baseName %in% imu)
```


The filtered list of files that we are interested in

```{r}
list_files_sub <- df_list_sub$full_path
```


## First pass 

> SKIP for now as we used bash to find files that has immuno 

In the first pass we want to read all the files and find studies with immuno drugs

```{r,eval=FALSE}

list_1 <- list()

# Run the workflow
for (i in seq_along(list_files_sub)){
  
  # Read file
  df <- read_xml(list_files_sub[i])
  
# ------------------ #
# Extracting study ID (clinical trials gov study ID)
# ------------------ #
xml_study_id <- 
  xml_text(
    xml_find_all(df, 
                 xpath = "//nct_id"))

#print(xml_study_id)
# ------------------ #
# Extracting study status (has the study been completed)
# ------------------ #
xml_study_status <- 
  xml_text(
    xml_find_all(df, 
                 xpath = "//overall_status"))

# ------------------ #
# Extracting if the study reported the results (some studies don't report the results)
# ------------------ #
xml_result_status <- 
  xml_text(
    xml_find_all(df, 
                 xpath = "//clinical_results"))


# ------------------ #
# Extracting condition (cancer type)
# ------------------ #
xml_cond <- 
  xml_text(
    xml_find_all(df, 
                 xpath = "//condition"))

# ------------------ #
# Extracting drug given in each arm of the study (intervention type)
# ------------------ #
xml_intervention <- 
  xml_text(
    xml_find_all(df, 
                 xpath = "//intervention_name"))

# ------------------ #
# Extracting the study arm (how many variations to the study, different dose / drugs)
# ------------------ #
# Number of arm
xml_arm <- 
  xml_text(
    xml_find_all(df, 
                 xpath = "//number_of_arms"))

# Title / name of arm
xml_arm_name <- 
  xml_text(
    xml_find_all(df, 
                 xpath = "//arm_group/arm_group_label"))

# Name of group based on clinical results
xml_clin_group <- 
  xml_text(
    xml_find_all(df, 
                 xpath = "//clinical_results/participant_flow/group_list/group/title"))

# Extract the reported event arm
ser_event_armTitle <- 
  xml_text(
    xml_find_all(df, 
                 xpath = "//reported_events/group_list/group/title"))

# ----------------------------- #
# --- Generating main DF ---- #
# ----------------------------- #

# Return required info, as some columns may be empty, we use the if_else function
# Df 1 - Study info 
df_study_info <- data.frame(
  
  study_id = 
    if_else(length(xml_study_id) >=1, 
            xml_study_id,"none",missing = NULL),
  study_status = 
    if_else(length(xml_study_status) >=1, 
            xml_study_status,"none",missing = NULL),
  #result_status =
  #  if_else(length(xml_result_status) >=1, 
  #          paste(xml_result_status, collapse = "_"),"none", missing = NULL),
  intervention = 
    if_else(length(xml_intervention) >=1, 
            paste(xml_intervention, collapse = "_"),"none", missing = NULL),
  condition = 
    if_else(length(xml_cond) >=1, 
            paste(xml_cond, collapse = "_"),"none", missing = NULL),
  arm_num = 
    if_else(length(xml_arm) >=1, 
            paste(xml_arm, collapse = "_"),"none", missing = NULL),
  arm_title = 
    if_else(length(xml_arm_name) >=1, 
            paste(xml_arm_name, collapse = "_"),"none", missing = NULL),
  clin_group =
        if_else(length(xml_clin_group) >=1, 
            paste(xml_clin_group, collapse = "_"),"none", missing = NULL),
  repEvent_title =
        if_else(length(ser_event_armTitle) >=1, 
            paste(ser_event_armTitle, collapse = "_"),"none", missing = NULL)
  )

list_1[[i]] <- df_study_info
}

```

Once we extracted the data, we compile everything into one big DF so we search for immuno drugs

```{r, eval=FALSE}
df_study_info <- do.call(rbind, list_1)
tt <- df_study_info[1:100,]
```

We simply search all the files in the data frame for any of the below immunotheraphy drugs

```{r,eval=FALSE}
# Does not work for some reason, need to debug
df_study_info_sub <-  df_study_info %>% 
  filter_all(any_vars(str_detect(., 
  "(?i)Ipilimumab|Yervoy|BMS-734016|MDX-010|MDX-101|
  Nivolumab|Opdivo|ONO-4538|BMS-936558|MDX1106|
  Pembrolizumab|Keytruda|MK-3475|lambrolizumab|
  Atezolizumab|Tecentriq|MPDL3280A|RG7446
  Durvalumab|Imfinzi|MEDI4736|MEDI-4736|
  Avelumab|Bavencio|MSB0010718C|
  Cemiplimab|Libtayo|REGN-2810|REGN2810|cemiplimab-rwl|
  Relatlimab|BMS-986016|relatlimab-rmbw|
  Dostarlimab|Jemperli|TSR-042|WBP-285|dostarlimab-gxly")))

# Also only select COMPLETED study
df_study_info_sub <- df_study_info_sub %>%
  dplyr::filter(study_status == "Completed")

```



## Second pass 


Write a loop / function to process the files as we did in objective 1. 

```{r}

list_1 <- list()
list_2 <- list()

# Run the workflow
for (i in seq_along(list_files_sub)){
  
  # Read file
  df <- read_xml(list_files_sub[i])
  
  # Tried to add a line to skip if the files does not have serious_events, cant seem to wget this to work
  #if(length(xml_find_all(df, xpath = "//reported_events/serious_events")) >=1 )  {
  

# ------------------ #
# Extracting study ID (clinical trials gov study ID)
# ------------------ #
xml_study_id <- 
  xml_text(
    xml_find_all(df, 
                 xpath = "//nct_id"))

#print(xml_study_id)
# ------------------ #
# Extracting study status (has the study been completed)
# ------------------ #
xml_study_status <- 
  xml_text(
    xml_find_all(df, 
                 xpath = "//overall_status"))

# ------------------ #
# Extracting if the study reported the results (some studies don't report the results)
# ------------------ #
xml_result_status <- 
  xml_text(
    xml_find_all(df, 
                 xpath = "//clinical_results"))


# ------------------ #
# Extracting condition (cancer type)
# ------------------ #
xml_cond <- 
  xml_text(
    xml_find_all(df, 
                 xpath = "//condition"))

# ------------------ #
# Extracting drug given in each arm of the study (intervention type)
# ------------------ #
xml_intervention <- 
  xml_text(
    xml_find_all(df, 
                 xpath = "//intervention_name"))

# ------------------ #
# Extracting the study arm (how many variations to the study, different dose / drugs)
# ------------------ #
# Number of arm
xml_arm <- 
  xml_text(
    xml_find_all(df, 
                 xpath = "//number_of_arms"))

# Title / name of arm
xml_arm_name <- 
  xml_text(
    xml_find_all(df, 
                 xpath = "//arm_group/arm_group_label"))

# Name of group based on clinical results
xml_clin_group <- 
  xml_text(
    xml_find_all(df, 
                 xpath = "//clinical_results/participant_flow/group_list/group/title"))
xml_clin_group <- xml_clin_group[!str_detect(xml_clin_group, "Not Enrolled")] # NCT01295827 listed the arm of patients not entorlled but did not list adverse effect, causing error
#xml_clin_group <- case_when(xml_clin_group == "Arm S: SUN + NIV2" ~ "Arm S: SUN + NIV",
#                            xml_clin_group == "Arm S: SUN + NIV5" ~ "Arm S: SUN + NIV",
#                            TRUE ~ xml_clin_group) # NCT01472081 combined the results of two arm
xml_clin_group <- unique(xml_clin_group) 

# ------------------ #
# Extracting serious events results
# ------------------ #
# Extract the Main text
ser_event_title <- 
  xml_text(
    xml_find_all(df, 
                 xpath = "//reported_events/serious_events/category_list/category/title"))

# Extract the Sub title text
ser_event_subTitle <- 
  xml_text(
    xml_find_all(df, 
                 xpath = "//reported_events/serious_events/category_list/category/event_list/event/sub_title"))

tmp_id <- c(xml_study_id,  ser_event_subTitle)

# NCT01497808 did not list one category of serious adverse event, need to manually add
if(identical(tmp_id, NCT01497808_ser_event_subTitle) == "TRUE"){
  ser_event_subTitle <-  NCT01497808_new
}
else{
  ser_event_subTitle <- ser_event_subTitle
}




# Extract the reported event arm
ser_event_armTitle <- 
  xml_text(
    xml_find_all(df, 
                 xpath = "//reported_events/group_list/group/title"))

# ------------------ #
# Extracting number of cases with severe irAE 
# ------------------ #
# Find the path which has the irAE event numbers
dat_path <- df %>% xml_find_all( '//counts') %>% xml_path # 42 lines
# Limit to serious effects
dat_path <- dat_path[grep("serious_events", dat_path) ]

# For all of the path found, extract the attributes using a loop
group_id <- vector()
num_case <- vector()
tot_case <- vector()


for (x in seq_along(dat_path)){
  ser_event_num <- 
    xml_attrs(xml_find_all(df, xpath = dat_path[x]))
  
  un_event <- unlist(ser_event_num)
  
  group_id[x] <- un_event["group_id"]
  num_case[x] <- un_event["subjects_affected"]
  tot_case[x] <- un_event["subjects_at_risk"]

}
#}

df_out <- data.frame(
  group_id = group_id,
  num_case = num_case,
  tot_case = tot_case)

# ----------------------------- #
# --- Generating main DF ---- #
# ----------------------------- #

# Return required info, as some columns may be empty, we use the if_else function
# Df 1 - Study info 
df_study_info <- data.frame(
  
  study_id = 
    if_else(length(xml_study_id) >=1, 
            xml_study_id,"none",missing = NULL),
  study_status = 
    if_else(length(xml_study_status) >=1, 
            xml_study_status,"none",missing = NULL),
  #result_status =
  #  if_else(length(xml_result_status) >=1, 
  #          paste(xml_result_status, collapse = "_"),"none", missing = NULL),
  intervention = 
    if_else(length(xml_intervention) >=1, 
            paste(xml_intervention, collapse = "_"),"none", missing = NULL),
  condition = 
    if_else(length(xml_cond) >=1, 
            paste(xml_cond, collapse = "_"),"none", missing = NULL),
  arm_num = 
    if_else(length(xml_arm) >=1, 
            paste(xml_arm, collapse = "_"),"none", missing = NULL),
  arm_title = 
    if_else(length(xml_arm_name) >=1, 
            paste(xml_arm_name, collapse = "_"),"none", missing = NULL),
  clin_group =
        if_else(length(xml_clin_group) >=1, 
            paste(xml_clin_group, collapse = "_"),"none", missing = NULL),
  repEvent_title =
        if_else(length(ser_event_armTitle) >=1, 
            paste(ser_event_armTitle, collapse = "_"),"none", missing = NULL)
  )


# Df 2- serious event numbers
# Some studies list fewer arm than results (eg - NCT01585987.xml), in cases like this we get the max arm as defined by the study 
max_arm <- length(ser_event_armTitle)
#max_arm <- length(xml_clin_group)
# As arm is listed as E1, E2 E3 ....etc, we generate Ex according to max arm number
e_arm <- paste0("E", 1:max_arm)
# Filter event to only these category
df_out <- df_out %>%
  dplyr::filter(group_id %in% e_arm)

# Some studies have multiple arms, resulting in each irAE listed multiple times. Find the number of arms. 
# Some studies list more arm than results (eg -  NCT00112580), in cases like this we reduce the number of arm
int_fac <- min(c(length(unique(df_out$group_id))), max_arm) # Take the minimum number
new_group <- vector()
if(int_fac == max_arm){
  new_group <- ser_event_armTitle
}else{
  new_group <- paste("first_group_only", ser_event_armTitle[1], sep=" ")
}

df_ser_number <- df_out %>%
  dplyr::mutate(study_id = xml_study_id,
                arm_title = rep(new_group, times=length(ser_event_subTitle)),
                serEvent = rep(ser_event_subTitle, each=int_fac))

  
list_1[[i]] <- df_study_info
list_2[[i]] <- df_ser_number

}

```

Save the list files into data frame 

Once we extracted the data, we compile everything into one big DF.

```{r}
df_study_info <- do.call(rbind, list_1)
df_av_eff <- do.call(rbind, list_2)
```

Save the file for future analysis

```{r}
save(df_study_info,df_av_eff,
     file=file.path(wd$dataP, "DF_ClinGov_immuno.Rdata"))
```




# References

Reading XML 
https://www.delftstack.com/howto/r/read-xml-in-r/
https://medium.com/geekculture/reading-xml-files-in-r-3122c3a2a8d9
https://stackoverflow.com/questions/52065772/extracting-attributes-from-xml-file
https://appsilon.com/r-xml/


XML to df https://stackoverflow.com/questions/17198658/how-to-parse-an-xml-file-to-an-r-data-frame
Nested list 
https://stackoverflow.com/questions/50367666/extract-contents-from-a-nested-list-in-r
https://stackoverflow.com/questions/63786411/how-to-unnest-wider-with-loop-over-all-the-columns-containing-lists
https://community.rstudio.com/t/extracting-from-nested-list/17016


